---
title: "Data_cleaning"
output: html_document
date: "2024-06-12"
---



This file includes the process to obtain the final dataset that we use for monthly analysis. The data cleaning process involves sourcing data from different resources and arranging it to match our new boundaries. To understand how we obtained the new boundaries, please refer to shapefile_new_from_2004.rmd.

The data processing in this file covers dengue cases, population, weather variables from ERA5 and weather station, preventive factors, and socio-economic factors.


##Dengue cases
```{r,warning=FALSE }
library(stringr)
library(ggplot2)
library(readr)
library(ggplot2)
library(spdep)
library(sp)
library(MASS)
library(surveillance)
library(tidyr)
library(splines)
library(tidyverse)
library(tsibble)
library(sf)
library(scales)
library(RColorBrewer)
library(viridis)
library(readxl)
library(dplyr)

Dengue<- read_excel("C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/HPC_District_monthly_update_lag/Raw_data/20231130_ED_MDR Dengue district data_2000-2023_original modification_updated 20May24.xlsx", sheet=1)

Dengue_filtered_data <- Dengue[Dengue$year >= 2004, ]


# Count missing values
sum(is.na(Dengue_filtered_data$Dengue))
sum(is.na(Dengue_filtered_data$SevereDHF))

Dengue_filtered_data <- Dengue_filtered_data %>%
  arrange(district, province, year, month)

# Define a function to fill missing values with the average from the same month in preceding years
fill_missing_with_average <- function(data, column) {
  data %>%
    group_by(district, province, month) %>%
    mutate({{column}} := ifelse(is.na({{column}}), mean({{column}}[year %in% (max(year) - 1):(min(year) + 1)], na.rm = TRUE), {{column}})) %>%
    ungroup()
}

# Apply the function to the columns "Dengue" and "SevereDHF"
Dengue_filtered_data <- Dengue_filtered_data %>%
  fill_missing_with_average(Dengue) %>%
  fill_missing_with_average(SevereDHF)

# Round up the filled values
Dengue_filtered_data <- Dengue_filtered_data %>%
  mutate(
    Dengue = ceiling(Dengue),
    SevereDHF = ceiling(SevereDHF)
  )

# Count the number of remaining missing values
missing_dengue <- sum(is.na(Dengue_filtered_data$Dengue))
missing_severe_dhf <- sum(is.na(Dengue_filtered_data$SevereDHF))

# Print the result
cat("Missing values in Dengue:", missing_dengue, "\n")
cat("Missing values in SevereDHF:", missing_severe_dhf, "\n")

# Match the boundaries
Dengue_filtered_data <- Dengue_filtered_data %>%
  mutate(district = toupper(district))

# Group and mutate the data to standardize district names
Dengue_filtered_data_summarized <- Dengue_filtered_data %>%
  group_by(province, district, year, month) %>%
  mutate(
    district = case_when(
      (district %in% c("VINH LOI", "HOA BINH")) & (province == "BAC LIEU") ~ "VINH LOI",
      (district %in% c("MO CAY","MO CAY NAM", "MO CAY BAC")) & (province == "BEN TRE") ~ "MO CAY",
      (district %in% c("CO DO", "THOI LAI")) & (province == "CAN THO") ~ "CO DO",
      (district %in% c("HONG NGU DISTRICT", "HONG NGU CITY" )) & (province == "DONG THAP") ~ "HONG NGU",
      (district %in% c("NGA BAY CITY", "PHUNG HIEP")) & (province == "HAU GIANG") ~ "NGA BAY",
      (district %in% c("LONG MY", "LONG MY TOWN")) & (province == "HAU GIANG") ~ "LONG MY",
      (district %in% c("AN MINH", "U MINH THUONG", "AN BIEN", "VINH THUAN")) & (province == "KIEN GIANG") ~ "AN MINH",
      (district %in% c("KIEN LUONG", "GIANG THANH")) & (province == "KIEN GIANG") ~ "KIEN LUONG",
      (district %in% c("KIEN TUONG", "MOC HOA")) & (province == "LONG AN") ~ "MOC HOA",
      (district %in% c("LONG PHU", "MY XUYEN", "TRAN DE")) & (province == "SOC TRANG") ~ "LONG PHU",
      (district %in% c("CHAU THANH", "MY TU")) & (province == "SOC TRANG") ~ "MY TU",
      (district %in% c("CAI LAY", "CAI LAY CITY", "CAI LAY DISTRICT")) & (province == "TIEN GIANG") ~ "CAI LAY",
      (district %in% c("GO CONG TAY", "GO CONG DONG", "TAN PHU DONG")) & (province == "TIEN GIANG") ~ "GO CONG DONG",
      (district %in% c("DUYEN HAI DISTRICT", "DUYEN HAI", "DUYEN HAI TOWN", "TRA CU")) & (province == "TRA VINH") ~ "TRA CU",
      (district %in% c("BINH MINH", "BINH TAN")) & (province == "VINH LONG") ~ "BINH MINH",
      TRUE ~ district
    )
  ) %>%
  summarise(Dengue = sum(Dengue, na.rm = TRUE), SevereDHF = sum(SevereDHF, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(district = ifelse(district == "RACH GIA CI TY", "RACH GIA CITY", district))

# Read the shapefile
MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")

# Convert district and province names to uppercase
MDR_NEW <- MDR_NEW %>%
  mutate(
    VARNAME = toupper(VARNAME),
    NAME_En = toupper(NAME_En)
  )

# Adjust district names in MDR_NEW
MDR_NEW <- MDR_NEW %>%
  mutate(VARNAME = gsub("CAM", "CA", VARNAME)) 


Dengue_filtered_data_summarized <- Dengue_filtered_data_summarized %>%
  mutate(
    district = toupper(district),
    province = toupper(province)
  )


# Join data with shapefile
MDR_Dengue_map_pre <- Dengue_filtered_data_summarized %>%
  left_join(MDR_NEW, by = c("district" = "VARNAME", "province" = "NAME_En"))

# Select required columns
selected_columns <- c(
  "province", "district", "year", "month", "Dengue", "SevereDHF",
  "NAME_1", "NAME_2", "TYPE_2", "ENGTYPE", "ID"
)

MDR_Dengue_map_pre <- MDR_Dengue_map_pre %>%
  dplyr::select(all_of(selected_columns))

MDR_Dengue_map_Final <- st_drop_geometry(MDR_Dengue_map_pre)
dim(MDR_Dengue_map_Final)
# Plotting for Bac Lieu
BAC_LIEU_N <- MDR_Dengue_map_Final %>%
  filter(district == 'BAC LIEU')

ggplot(BAC_LIEU_N, aes(x = month, y = Dengue, group = year, color = factor(year))) +
  geom_line() +
  labs(title = "Monthly Dengue Cases in Bac Lieu",
       x = "Month",
       y = "Number of Dengue Cases",
       color = "Year") +
  theme_minimal()

# Create a date column
BAC_LIEU_N <- BAC_LIEU_N %>%
  mutate(date = as.Date(paste(year, month, "1", sep = "-")))

# Create a time series plot
ggplot(BAC_LIEU_N, aes(x = date, y = Dengue)) +
  geom_line() +
  labs(title = "Time Series of Dengue Cases in Bac Lieu",
       x = "Date",
       y = "Number of Dengue Cases") +
  theme_minimal()
```



## Populaion data

```{r setup, include=FALSE,warning=False}
file_path <- "C:\\Users\\uqwareed\\OneDrive - The University of Queensland\\Re District-level data_DUNG\\MDR_pop_district-Eng version- LT.xlsx"

df <- read_excel(file_path, sheet = 1)
df<- df[,c(3,5,6,8,9)]
df<- df[df$Year >=2004,]
df$District <- str_to_title(df$District)
df$Province<- str_to_title(df$Province)
df$prov_dis <- paste0(df$District,"", df$Province)

pop_filtered_data_summarized <- df %>%  group_by( Year, District,Province)%>%
  mutate(
        District = 
      ifelse(
        (  District %in% c("Mo Cay Nam", "Mo Cay Bac")) &  Province %in% 'Ben Tre',
       "Mo Cay",
        ifelse(
          (  District %in%c("Co Do","Thoi Lai")) &  Province %in% 'Can Tho',
         "Co Do",
          ifelse(
            (  District %in% c("Hong Ngu City", "Hong Ngu Town"))  &  Province %in% 'Dong Thap',
           "Hong Ngu",
            ifelse(
              (  District %in% c("Nga Bay City", "Nga Bay Town","Phung Hiep")) &  Province %in% "Hau Giang",
              "Nga Bay",
              ifelse(
                (  District %in% c("Long My", "Long My Town")) &  Province %in% "Hau Giang" ,
                "Long My",
                ifelse(
                  (  District %in%  c("An Minh", "Vinh Thuan",  "U Minh Thuong" ,"An Bien"  )) &  Province%in% "Kien Giang",
                  "An Minh",
                  ifelse(
                    (  District %in% c("Kien Luong", "Giang Thanh")) &  Province%in% "Kien Giang",
                   "Kien Luong",
                    ifelse(
                      (  District %in% c("Moc Hoa" , "Kien Tuong Town")) &  Province %in% "Long An",
                      "Moc Hoa" ,
                      ifelse(
                        (  District %in% c("Long Phu"  , "My Xuyen", "Tran De")) &  Province %in% "Soc Trang",
                        "Long Phu"  ,
                        ifelse(
                        (  District %in%c("Chau Thanh", "My Tu" )) &  Province %in% "Soc Trang",
                        "My Tu",
                        ifelse(
                          (  District %in% c("Cai Lay", "Cai Lay City" ) )&  Province %in% "Tien Giang",
                          "Cai Lay",
                          ifelse(
                            (  District %in% c("Go Cong Dong", "Go Cong Tay", "Tan Phu Dong" )) &  Province %in% "Tien Giang" ,
                            "Go Cong Dong",
                            ifelse(
                              (  District %in% c("Duyen Hai", "Duyen Hai Town","Tra Cu" )) &  Province %in% "Tra Vinh" ,
                              "Tra Cu" ,
                              ifelse(
                                (  District %in%  c("Binh Minh Town", "Binh Tan" ) ) &  Province %in% "Vinh Long",
                               "Binh Minh",
                               ifelse(
       District %in% c("Vinh Loi", "Hoa Binh") &  Province %in% "Bac Lieu",
      "Vinh Loi",   District
                              )
      

                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  ) )%>%
 
  summarise(
    total_population = sum(Total_population,na.rm = TRUE),Population_density=sum(Population_density,na.rm=TRUE))%>%
  ungroup()

pop_filtered_data_summarized$Province_District<-paste0(pop_filtered_data_summarized$Province,"_",pop_filtered_data_summarized$District)




# Create a new data frame to store the interpolated data
#repeated the same value for each month 


# Assuming filled_data is your original data frame
# Create a new data frame with the 'Month' column
expanded_data <- expand.grid(Month = 1:12, District = unique(pop_filtered_data_summarized$District), Province = unique(pop_filtered_data_summarized$Province))

# Merge with filled_data to get the required columns
expanded_data <- merge(expanded_data, pop_filtered_data_summarized, by = c("District", "Province"))

# Use tidyr's complete function to create all combinations of Year, Month, District, and Province
expanded_data <- arrange(complete(expanded_data, Year, Month, nesting(District, Province)))
expanded_data <- expanded_data %>%
  arrange(Year, Month)
# Print the expanded_data
print(expanded_data)

data_2021 <- expanded_data %>%
  filter(Year == 2021 & District == "Cao Lanh City" & Province == "Dong Thap") 

# Check the extracted 2021 data
print(data_2021)

# Fill in the missing 2022 data with the 2021 data
expanded_data <- expanded_data %>%
  mutate(total_population = ifelse(Year == 2022 & District == "Cao Lanh City" & Province == "Dong Thap",
                                   data_2021$total_population[match(Month, data_2021$Month)],
                                   total_population),
         Population_density = ifelse(Year == 2022 & District == "Cao Lanh City" & Province == "Dong Thap",
                                     data_2021$Population_density[match(Month, data_2021$Month)],
                                     Population_density))


na_rows <- expanded_data[is.na(expanded_data$total_population), ]

##fill the one year missing pop data for CAO LANH CITY in 2022 to be like in 2021

#Join it with the new shape file


MDR_NEW <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")


MDR_NEW$VARNAME<- toupper(MDR_NEW$VARNAME)
expanded_data$District<- toupper(expanded_data$District)

MDR_NEW$NAME_En<- toupper(MDR_NEW$NAME_En)
expanded_data$Province<- toupper(expanded_data$Province)

setdiff(MDR_NEW$VARNAME,expanded_data$District)
setdiff(expanded_data$District,MDR_NEW$VARNAME)
# Create a mapping between old names and new names
district_mapping <- c(  "CAO LANH" ="CAO LANH DISTRICT",
                      "GIA RAI TOWN" = "GIA RAI" ,
                      "RACH GIA"= "RACH GIA CITY",
                      "NGA NAM TOWN" ="NGA NAM" ,
                     "SOC TRANG" = "SOC TRANG CITY" ,
                      "TRA VINH" = "TRA VINH CITY" ,
                      "VINH LONG" = "VINH LONG CITY",
                     "TAN CHAU TOWN"  ="TAN CHAU")

# Replace the names in filled_data using the mapping
expanded_data <- expanded_data%>%
  mutate(
   District  = ifelse(District  %in% names(district_mapping), district_mapping[District ], District )
  )

##Fill the missing values 


setdiff(MDR_NEW$VARNAME,expanded_data$District)
setdiff(expanded_data$District,MDR_NEW$VARNAME)

expanded_data$Province_District<- paste0(expanded_data$Province,"_",expanded_data$District)
MDR_NEW$province_district<- toupper(paste(MDR_NEW$NAME_En, MDR_NEW$VARNAME, sep = "_"))

setdiff(MDR_NEW$province_district,expanded_data$Province_District)
setdiff(expanded_data$Province_District,MDR_NEW$province_district)

# Assuming 'expanded_data' is your tibble
MDR_NEW<- MDR_NEW %>%
  mutate(VARNAME = gsub("CAM", "CA", VARNAME))

MDR_NEW<- MDR_NEW %>%
  mutate(province_district = gsub("CAM", "CA", province_district))

# View the updated tibble
print(expanded_data)


 matching_district <- (MDR_NEW$province_district %in% expanded_data$Province_District )
  matching_district

# Display the matching values
matched_names <- unique(MDR_NEW$province_district[matching_district])

matched_names

Non_matching<- unique(MDR_NEW$province_district[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)


MDR_pop <- left_join(MDR_NEW, expanded_data, by = c("province_district"="Province_District"))

dim(MDR_pop)


selected_columns <- c(
   "NAME_1", "NAME_2", "TYPE_2","Year","Month","VARNAME",
  "ENGTYPE",  "total_population","District",     "Province","Population_density"
  
)

Final_set<- inner_join(MDR_Dengue_map_Final, MDR_pop, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="Year","month"="Month","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))


dim(Final_set)

# Assuming your existing data frame is named existing_data_frame
# Assuming you want to select specific columns based on the provided names

# Columns to select
selected_columns <- c(
   "province", "district", "year", "month", "Dengue", "SevereDHF",
   "NAME_1", "NAME_2", "TYPE_2","Population_density",
  "ENGTYPE",   "total_population"
)

# Create the new data frame
Final_set_subset <- Final_set %>%
  dplyr::select(all_of(selected_columns))


pop_Dengue_Final_set_subset<- Final_set_subset
# Print the final subset
dim(pop_Dengue_Final_set_subset)

# Check for rows with any NA values
na_rows <- pop_Dengue_Final_set_subset[!complete.cases(pop_Dengue_Final_set_subset), ]

# Display the rows with NA values, if any
if (nrow(na_rows) > 0) {
  print("Rows with NA values:")
  print(na_rows)
} else {
  print("No rows with NA values found.")
}

```




##Socioeconomic variables
The NIHE team has provided socioeconomic data obtained from the census.

```{r}
file_path <- "C:\\Users\\uqwareed\\OneDrive - The University of Queensland\\Re District-level data_DUNG\\output_file-LT_revised-w-note_v4.xlsx"
df <- read_excel(file_path, sheet = 2)
df<- df[df$Year>=2004,]

#The provided code snippet is intended for filling in missing data concerning Poverty Rate, Access to Hygienic Water, Access to Hygienic #Toilets, Monthly Average Income Per Capital, Total Passenger, and Urbanization Rate. since the data are cooolected for every two years..


# Sort the data frame by Province and Year
df <- df %>% arrange(Province, Year)

# Group by Province and apply linear interpolation to Urbanization_Rate
df <- df %>%
  group_by(Province) %>%
  mutate(Urbanization_Rate = zoo::na.approx(Urbanization_Rate, na.rm = FALSE))

# Fill missing values with the average of the previous and next year for the same province
df <- df %>%
  group_by(Province) %>%
  mutate(Urbanization_Rate = ifelse(is.na(Urbanization_Rate),
                                    (lag(Urbanization_Rate) + lead(Urbanization_Rate, default = last(Urbanization_Rate))) / 2,
                                    Urbanization_Rate))



# Group by Province and apply linear interpolation to Urbanization_Rate
df <- df %>%
  arrange(Province, Year) %>%
  group_by(Province) %>%
  mutate(
    Poverty_Rate = zoo::na.approx(Poverty_Rate, na.rm = FALSE),
    Hygienic_Water_Access = zoo::na.approx(Hygienic_Water_Access, na.rm = FALSE),
    Hygienic_Toilet_Access = zoo::na.approx(Hygienic_Toilet_Access, na.rm = FALSE),
    Monthly_Average_Income_Percapita = zoo::na.approx(Monthly_Average_Income_Percapita, na.rm = FALSE),
    Total_Passenger = zoo::na.approx(Total_Passenger, na.rm = FALSE),
    Urbanization_Rate = zoo::na.approx(Urbanization_Rate, na.rm = FALSE)
  )


# Fill missing values with the average of the previous and next year for the same province
df <- df %>%
  arrange(Province, Year) %>%
  group_by(Province) %>%
  mutate(
    Urbanization_Rate = ifelse(is.na(Urbanization_Rate),
                               (lag(Urbanization_Rate) + lead(Urbanization_Rate, default = last(Urbanization_Rate))) / 2,
                               Urbanization_Rate),
    
    Poverty_Rate = ifelse(is.na(Poverty_Rate),
                          (lag(Poverty_Rate) + lead(Poverty_Rate, default = last(Poverty_Rate))) / 2,
                          Poverty_Rate),
    
    Hygienic_Water_Access = ifelse(is.na(Hygienic_Water_Access),
                                   (lag(Hygienic_Water_Access) + lead(Hygienic_Water_Access, default = last(Hygienic_Water_Access))) / 2,
                                   Hygienic_Water_Access),
    
    Hygienic_Toilet_Access = ifelse(is.na(Hygienic_Toilet_Access),
                                    (lag(Hygienic_Toilet_Access) + lead(Hygienic_Toilet_Access, default = last(Hygienic_Toilet_Access))) / 2,
                                    Hygienic_Toilet_Access),
    
    Monthly_Average_Income_Percapita = ifelse(is.na(Monthly_Average_Income_Percapita),
                                              (lag(Monthly_Average_Income_Percapita) + lead(Monthly_Average_Income_Percapita, default = last(Monthly_Average_Income_Percapita))) / 2,
                                              Monthly_Average_Income_Percapita),
    
    Total_Passenger = ifelse(is.na(Total_Passenger),
                             (lag(Total_Passenger) + lead(Total_Passenger, default = last(Total_Passenger))) / 2,
                             Total_Passenger)
  )





Outmigration_Rate_df <- df[, c(1, 2, 11)]

# Fill missing values for 2004 with values from 2005 for each province
Outmigration_Rate_df <- Outmigration_Rate_df %>%
  group_by(Province) %>%
  fill(Outmigration_Rate, .direction = "downup")

# Print the updated data frame
print(Outmigration_Rate_df)

df$Outmigration_Rate<- Outmigration_Rate_df$Outmigration_Rate


# Assuming your data frame is named df
Inmigration_Rate_df <- df[, c(1, 3, 11)]

# Assuming your data frame is named df
s_filled <- Inmigration_Rate_df %>%
  group_by(Province) %>%
  mutate(Inmigration_Rate = ifelse(is.na(Inmigration_Rate) & Year == 2004, lead(Inmigration_Rate), Inmigration_Rate),
         Inmigration_Rate = ifelse(is.na(Inmigration_Rate) & Year == 2006, (lag(Inmigration_Rate) + lead(Inmigration_Rate)) / 2, Inmigration_Rate))

# View the filled data frame
#View(s_filled)

df$Inmigration_Rate<- s_filled$Inmigration_Rate


# Assuming your data frame is named df
NetImmigration_Rate <- df[, c(1, 4, 11)]


# Assuming your data frame is named df
s2_filled <- NetImmigration_Rate %>%
  group_by(Province) %>%
  mutate(NetImmigration_Rate = ifelse(is.na(NetImmigration_Rate) & Year == 2004, lead(NetImmigration_Rate), NetImmigration_Rate),
         NetImmigration_Rate = ifelse(is.na(NetImmigration_Rate) & Year == 2006, (lag(NetImmigration_Rate) + lead(NetImmigration_Rate)) / 2, NetImmigration_Rate))



df$NetImmigration_Rate<- s2_filled$NetImmigration_Rate
s<- df %>%
  group_by(Province) %>%
  mutate(NetImmigration_Rate = ifelse(is.na(NetImmigration_Rate), (df$Inmigration_Rate)-(df$Outmigration_Rate), NetImmigration_Rate))

df$NetImmigration_Rate<- s$NetImmigration_Rate


#Fill the poverty rate for the years 2004 and 2005 for each province using the linear interpolation.


# Assuming your data frame is named df
d <- df[, c(1, 5, 11)]

# Identify the rows with missing values for 2004 and 2005
missing_rows <- which(is.na(d$Poverty_Rate) & (d$Year == 2004 ))

# Loop through each Province and fill missing values using linear model predictions
for (province in unique(d$Province)) {
  # Subset data for the current province
  province_data <- d[d$Province == province, ]
  
  # Create a linear model
  lm_model <- lm(Poverty_Rate ~ Year, data = province_data)
  
  # Predict missing values
  predictions <- predict(lm_model, newdata = data.frame(Year = c(2004)))
  
  # Fill in missing values for the current province
  d$Poverty_Rate[missing_rows] <- predictions
}

# View the updated data frame


##filling 2005 
missing_rows <- which(is.na(d$Poverty_Rate) & (d$Year == 2005))

# Loop through each Province and fill missing values using linear model predictions
for (province in unique(d$Province)) {
  # Subset data for the current province
  province_data <- d[d$Province == province, ]
  
  # Create a linear model
  lm_model <- lm(Poverty_Rate ~ Year, data = province_data)
  
  # Predict missing values
  predictions <- predict(lm_model, newdata = data.frame(Year = c(2005)))
  
  # Fill in missing values for the current province
  d$Poverty_Rate[missing_rows] <- predictions
}



df$Poverty_Rate<- d$Poverty_Rate

df <- df[, c(1:11)]

missing_values <- is.na(df)

# Summarize the count of missing values in each column
missing_count <- colSums(missing_values)

# Display the count of missing values for each column
print(missing_count)

##match it with dengue cases 
socio_dengue_pop<- left_join(pop_Dengue_Final_set_subset,df,by=c("province"="Province","year"="Year"))


```

##Weather variables from ERA5
ws_avg: daily average wind speed (m/s)
t2m_avg: daily average temperature (in deg K = deg C + 273.15)
rh_avg: daily average relative humidity (%)
 ws_max: daily maximum wind speed (m/s)
 t2m_max: daily maximum temperature (in deg K = deg C + 273.15)
 rh_max: daily maximum relative humidity (%)
 ws_min: daily minimum wind speed (m/s)
t2m_min: daily minimum temperature (in deg K = deg C + 273.15)
rh_min: daily minimum relative humidity (%)
 tp_accum: accumulated total precipitation (m)
 
 
 These data obtained from Van Github per request 

```{r}

# Set the directory path

##These data can be found in Van Github per request
directory_path <- "C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/metdata/ERA5/ERA5land_data"

# Get a list of all files in the directory with a .csv extension
files <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

##These lines will take long time to run 
# Initialize an empty data frame to store the results
result_df <- data.frame()
# Iterate through the list of files and read each CSV file
for (file in files) {
  # Extract year and month from the file name
  file_parts <- strsplit(basename(file), "\\.")[[1]]
  year_month <- substr(file_parts[1], 1, 6)
  year <- substr(year_month, 1, 4)
  month <- substr(year_month, 5, 6)

  # Read the CSV file
  df <- read_csv(file, col_types = cols(.default = "c"))
df1 <- apply(df[-c(1,2),2:ncol(df)], 2, function(x) as.numeric(as.character(x)))
  # Calculate the mean for each column
  means <- colMeans(df1, na.rm = TRUE)

  # Combine year, month, and mean into a data frame
  summary_data <- data.frame(
    Year = as.integer(year),
    Month = as.integer(month),ID= as.vector(as.character(df[1,2:ncol(df)])),
    Mean = means,names= as.vector(as.character(df[1,2:ncol(df)]))
  )

  # Append the summary data to the result data frame
  result_df <- bind_rows(result_df, summary_data)
}

# Print the resulting data frame
#print(result_df)
#saveRDS(result_df,'result_df.rds')


##Reshape the data set

result_df<- readRDS('C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/result_df.rds')
# Assuming result_df is your data frame
unique_names <- unique(result_df$names)


new_df <- result_df %>%
  group_by(Year, Month) %>%
  summarize(
    ws_avg = Mean[names == "ws_avg"],
    ws_max = Mean[names == "ws_max"],
    ws_min = Mean[names == "ws_min"],
    t2m_avg = Mean[names == "t2m_avg"],
    t2m_max = Mean[names == "t2m_max"],
    t2m_min = Mean[names == "t2m_min"],
    rh_avg = Mean[names == "rh_avg"],
    rh_max = Mean[names == "rh_max"],
    rh_min = Mean[names == "rh_min"],
    tp_accum = Mean[names == "tp_accum"]
  ) %>%
  ungroup()  # Remove grouping for the new data frame

print(new_df)

#add the ID
result_df <- new_df %>%
  mutate(ID = rep(0:133, length.out = n()))

# View the updated data frame
print(result_df)
tail(result_df)


#Match the ID with the  VARNAME_2 name

VARNAME_2_inf<- read.csv('C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/district_infomation.csv')

VARNAME_2_inf$X<- as.numeric(VARNAME_2_inf$X)
unique(VARNAME_2_inf$X)
result_df$ID <- as.numeric(result_df$ID)
unique(result_df$ID)

setdiff(unique(VARNAME_2_inf$X),unique(result_df$ID))

DAT<- left_join(result_df, VARNAME_2_inf,by=c('ID'='X'))

unique(paste(DAT$VARNAME_2, DAT$ID, sep = "_"))

unique(paste(DAT$VARNAME_2, DAT$NAME_1,DAT$NAME_2, sep = "_"))

#This shape file obtained form Van github 
MDR_2 <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re Districts in file_shapefile/VMD_districts_2022.shp")

DAT_MDR_2<- left_join(DAT,MDR_2,by=c("VARNAME_2"="VARNAME_2","NAME_1" ="NAME_1" ,"NAME_2"="NAME_2" ))

library(sf)

Full_data<- st_drop_geometry(DAT_MDR_2)

filtered_data <- Full_data[Full_data$Year >= 2004, ]

dim(filtered_data)
names(filtered_data)

filtered_data<-filtered_data[,-c(13,17,18,19,20,21,22,23,24,25,26,27,16)]
names(filtered_data)

# Assuming your data frame is named filtered_data.
filtered_data_summarized <- filtered_data %>%  group_by(NAME_1 ,NAME_2, Year, Month)%>%
  mutate(
       NAME_2 = 
      ifelse(
        ( NAME_2 %in% c("Mỏ Cày Nam", "Mỏ Cày Bắc")) & NAME_1 %in% "Bến Tre",
       "Mỏ Cày",
        ifelse(
          ( NAME_2 %in%c( "Cờ Đỏ","Thới Lai")) & NAME_1 %in% "Cần Thơ",
          "Cờ Đỏ",
          ifelse(
            ( NAME_2 %in% c("Hồng Ngự (Thành phố)", "Hồng Ngự"))  & NAME_1 %in% "Đồng Tháp",
           "Hồng Ngự",
            ifelse(
              ( NAME_2 %in% c("Ngã Bảy", "Phụng Hiệp")) & NAME_1 %in% "Hậu Giang",
              "Ngã Bảy",
              ifelse(
                ( NAME_2 %in% c("Long Mỹ", "Long Mỹ (Thị xã)")) & NAME_1 %in% "Hậu Giang" ,
                "Long Mỹ",
                ifelse(
                  ( NAME_2 %in%  c("An Minh", "Vĩnh Thuận", "U Minh Thượng" ,"An Biên")) & NAME_1%in% "Kiên Giang",
                  "An Minh",
                  ifelse(
                    ( NAME_2 %in% c("Kiên Lương", "Giang Thành")) & NAME_1%in% "Kiên Giang",
                    "Kiên Lương",
                    ifelse(
                      ( NAME_2 %in% c("Mộc Hóa", "Kiến Tường")) & NAME_1 %in% "Long An",
                      "Mộc Hóa",
                      ifelse(
                        ( NAME_2 %in% c("Long Phú", "Mỹ Xuyên", "Trần Đề" )) & NAME_1 %in% "Sóc Trăng",
                        "Long Phú",
                        ifelse(
                        ( NAME_2 %in%c("Châu Thành", "Mỹ Tú" )) & NAME_1 %in% "Sóc Trăng",
                        "Mỹ Tú",
                        ifelse(
                          ( NAME_2 %in% c("Cai Lậy", "Cai Lậy (Thị xã)" ) )& NAME_1 %in% "Tiền Giang",
                          "Cai Lậy",
                          ifelse(
                            ( NAME_2 %in% c("Gò Công Đông", "Gò Công Tây", "Tân Phú Đông" )) & NAME_1 %in% "Tiền Giang" ,
                            "Gò Công Đông",
                            ifelse(
                              ( NAME_2 %in% c("Duyên Hải", "Duyên Hải (Thị xã)","Trà Cú")) & NAME_1 %in% "Trà Vinh" ,
                              "Trà Cú",
                              ifelse(
                                ( NAME_2 %in%  c("Bình Minh", "Bình Tân") ) & NAME_1 %in% "Vĩnh Long",
                               "Bình Minh",
                               ifelse(
      NAME_2 %in% c("Vĩnh Lợi", "Hoà Bình") & NAME_1 %in% "Bạc Liêu",
      "Vĩnh Lợi",  NAME_2
                              )
                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  ) )%>%
 
  summarise(ws_avg = mean(ws_avg), ws_max = mean(ws_max), ws_min = mean(ws_min), t2m_avg = mean(t2m_avg), t2m_max = mean(t2m_max), t2m_min = mean(t2m_min), rh_avg = mean(rh_avg), rh_max = mean(rh_max),rh_min = mean(rh_min),tp_accum = mean(tp_accum)) %>%
  ungroup()


dim(filtered_data_summarized )


##Check the dimension to be similar to the dengue cases.
filtered_data_summarized_without_2023_year <- filtered_data_summarized %>%
  filter(!(Year == 2023))
MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")

dim(filtered_data_summarized_without_2023_year )
MDR_Dengue_map_pre <- left_join(MDR_NEW, filtered_data_summarized, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))

dim(MDR_Dengue_map_pre)

selected_columns <- c(
   "NAME_1", "NAME_2", "TYPE_2","Year","Month","VARNAME",
  "ENGTYPE",  "ws_avg", "ws_max", "ws_min", "t2m_avg",
  "t2m_max", "t2m_min", "rh_avg", "rh_max", "rh_min", "tp_accum"
)

# Create the new data frame
D <- MDR_Dengue_map_pre %>%
  dplyr::select(all_of(selected_columns))
D1<- st_drop_geometry(D)

 


MDR_weather_map <- left_join(MDR_NEW, filtered_data_summarized_without_2023_year, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))


MDR_weather_map$NAME_En<- toupper(MDR_weather_map$NAME_En)
MDR_weather_map$VARNAME<- toupper(MDR_weather_map$VARNAME)

Final_set<- inner_join(socio_dengue_pop, MDR_weather_map, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="Year","month"="Month","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))

dim(Final_set)

# Assuming your existing data frame is named existing_data_frame
# Assuming you want to select specific columns based on the provided names

# Columns to select
selected_columns <- c(
   "province", "district", "year", "month", "Dengue", "SevereDHF",
     "ws_avg", "ws_max", "ws_min", "t2m_avg",
  "t2m_max", "t2m_min", "rh_avg", "rh_max", "rh_min", "tp_accum",
  "total_population"  ,               "Outmigration_Rate",               
 "Inmigration_Rate"               ,  "NetImmigration_Rate" ,             "Poverty_Rate" ,                   
 "Hygienic_Water_Access"           , "Monthly_Average_Income_Percapita" ,"Total_Passenger",                 
 "Hygienic_Toilet_Access"       ,    "Urbanization_Rate"
)

# Create the new data frame
socio_dengue_pop_subset_ERA5_weather <- Final_set %>%
  dplyr::select(all_of(selected_columns))

```


##Weather data from weather station 

ws_avg: daily average wind speed (m/s)
t2m_avg: daily average temperature (in deg K = deg C + 273.15)
rh_avg: daily average relative humidity (%)
 ws_max: daily maximum wind speed (m/s)
 t2m_max: daily maximum temperature (in deg K = deg C + 273.15)
 rh_max: daily maximum relative humidity (%)
 ws_min: daily minimum wind speed (m/s)
t2m_min: daily minimum temperature (in deg K = deg C + 273.15)
rh_min: daily minimum relative humidity (%)
 tp_accum: accumulated total precipitation (m)


```{r}


# Set the directory path
directory_path <- "C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/metdata/IMHEN/04_data4edengue/v2"

# Get a list of all files in the directory with a .csv extension
files <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)


# Initialize an empty data frame to store the results
result_df <- data.frame()
# Iterate through the list of files and read each CSV file
for (file in files) {
  # Extract year and month from the file name
  file_parts <- strsplit(basename(file), "\\.")[[1]]
  year_month <- substr(file_parts[1], 1, 6)
  year <- substr(year_month, 1, 4)
  month <- substr(year_month, 5, 6)

  # Read the CSV file
  df <- read_csv(file, col_types = cols(.default = "c"))
df1 <- apply(df[-c(1,2),2:ncol(df)], 2, function(x) as.numeric(as.character(x)))
  # Calculate the mean for each column
  means <- colMeans(df1, na.rm = TRUE)

  # Combine year, month, and mean into a data frame
  summary_data <- data.frame(
    Year = as.integer(year),
    Month = as.integer(month),ID= as.vector(as.character(df[1,2:ncol(df)])),
    Mean = means,names= as.vector(as.character(df[1,2:ncol(df)]))
  )

  # Append the summary data to the result data frame
  result_df <- bind_rows(result_df, summary_data)
}

# Print the resulting data frame
#print(result_df)
#saveRDS(result_df,'result_df_weather_station_v3.rds')

result_df<- readRDS('C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/result_df_weather_station_v2.rds')
#result_df<- readRDS('result_df.rds')
# Assuming result_df is your data frame
unique_names <- unique(result_df$names)


new_df <- result_df[,-c(3)] %>%
  group_by(Year, Month) %>%
  summarize(
    t2m_avg = Mean[names == "t2m_avg"],
    t2m_max = Mean[names == "t2m_max"],
    t2m_min = Mean[names == "t2m_min"],
     tp_accum = Mean[names == "tp_accum"],
    rh_avg=Mean[names == "rh_avg"],
    rh_min=Mean[names == "rh_min"],
    ws_avg=Mean[names == "ws_avg"]
    
  ) %>%
  ungroup()  # Remove grouping for the new data frame


#add the ID
result_df <- new_df %>%
  mutate(ID = rep(0:133, length.out = n()))

# View the updated data frame
print(result_df)
tail(result_df)



VARNAME_2_inf<- read.csv('C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/district_infomation.csv')

VARNAME_2_inf$X<- as.numeric(VARNAME_2_inf$X)
unique(VARNAME_2_inf$X)
result_df$ID <- as.numeric(result_df$ID)
unique(result_df$ID)

setdiff(unique(VARNAME_2_inf$X),unique(result_df$ID))

DAT<- left_join(result_df, VARNAME_2_inf,by=c('ID'='X'))

unique(paste(DAT$VARNAME_2, DAT$ID, sep = "_"))

unique(paste(DAT$VARNAME_2, DAT$NAME_1,DAT$NAME_2, sep = "_"))


MDR_2 <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re Districts in file_shapefile/VMD_districts_2022.shp")

DAT_MDR_2<- left_join(DAT,MDR_2,by=c("VARNAME_2"="VARNAME_2","NAME_1" ="NAME_1" ,"NAME_2"="NAME_2" ))



library(sf)

Full_data<- st_drop_geometry(DAT_MDR_2)

filtered_data <- Full_data[Full_data$Year >= 2004, ]

dim(filtered_data)
names(filtered_data)

filtered_data<-filtered_data[,-c(13:ncol(filtered_data))]
names(filtered_data)


filtered_data_summarized <- filtered_data %>%  group_by(NAME_1 ,NAME_2, Year, Month)%>%
  mutate(
       NAME_2 = 
      ifelse(
        ( NAME_2 %in% c("Mỏ Cày Nam", "Mỏ Cày Bắc")) & NAME_1 %in% "Bến Tre",
       "Mỏ Cày",
        ifelse(
          ( NAME_2 %in%c( "Cờ Đỏ","Thới Lai")) & NAME_1 %in% "Cần Thơ",
          "Cờ Đỏ",
          ifelse(
            ( NAME_2 %in% c("Hồng Ngự (Thành phố)", "Hồng Ngự"))  & NAME_1 %in% "Đồng Tháp",
           "Hồng Ngự",
            ifelse(
              ( NAME_2 %in% c("Ngã Bảy", "Phụng Hiệp")) & NAME_1 %in% "Hậu Giang",
              "Ngã Bảy",
              ifelse(
                ( NAME_2 %in% c("Long Mỹ", "Long Mỹ (Thị xã)")) & NAME_1 %in% "Hậu Giang" ,
                "Long Mỹ",
                ifelse(
                  ( NAME_2 %in%  c("An Minh", "Vĩnh Thuận", "U Minh Thượng" ,"An Biên")) & NAME_1%in% "Kiên Giang",
                  "An Minh",
                  ifelse(
                    ( NAME_2 %in% c("Kiên Lương", "Giang Thành")) & NAME_1%in% "Kiên Giang",
                    "Kiên Lương",
                    ifelse(
                      ( NAME_2 %in% c("Mộc Hóa", "Kiến Tường")) & NAME_1 %in% "Long An",
                      "Mộc Hóa",
                      ifelse(
                        ( NAME_2 %in% c("Long Phú", "Mỹ Xuyên", "Trần Đề" )) & NAME_1 %in% "Sóc Trăng",
                        "Long Phú",
                        ifelse(
                        ( NAME_2 %in%c("Châu Thành", "Mỹ Tú" )) & NAME_1 %in% "Sóc Trăng",
                        "Mỹ Tú",
                        ifelse(
                          ( NAME_2 %in% c("Cai Lậy", "Cai Lậy (Thị xã)" ) )& NAME_1 %in% "Tiền Giang",
                          "Cai Lậy",
                          ifelse(
                            ( NAME_2 %in% c("Gò Công Đông", "Gò Công Tây", "Tân Phú Đông" )) & NAME_1 %in% "Tiền Giang" ,
                            "Gò Công Đông",
                            ifelse(
                              ( NAME_2 %in% c("Duyên Hải", "Duyên Hải (Thị xã)","Trà Cú")) & NAME_1 %in% "Trà Vinh" ,
                              "Trà Cú",
                              ifelse(
                                ( NAME_2 %in%  c("Bình Minh", "Bình Tân") ) & NAME_1 %in% "Vĩnh Long",
                               "Bình Minh",
                               ifelse(
      NAME_2 %in% c("Vĩnh Lợi", "Hoà Bình") & NAME_1 %in% "Bạc Liêu",
      "Vĩnh Lợi",  NAME_2
                              )
                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  ) )%>%
 
  summarise( t2m_avg = mean(t2m_avg), t2m_max = mean(t2m_max), t2m_min = mean(t2m_min),tp_accum=mean(tp_accum),
           rh_avg=mean(rh_avg),rh_min=mean(rh_min),ws_avg=mean(ws_avg)  ) %>%
  ungroup()


dim(filtered_data_summarized )


##Check the dimension to be similar to the dengue cases.
filtered_data_summarized_without_2023_year <- filtered_data_summarized %>%
  filter(!(Year == 2023))
MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries_Final.shp")

dim(filtered_data_summarized_without_2023_year )
MDR_Dengue_map_pre <- left_join(MDR_NEW, filtered_data_summarized, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))

dim(MDR_Dengue_map_pre)

selected_columns <- c(
   "NAME_1", "NAME_2", "TYPE_2","Year","Month","VARNAME",
  "ENGTYPE",   "t2m_avg", "tp_accum",
  "t2m_max", "t2m_min","rh_avg",   "rh_min" ,  "ws_avg"  
)

# Create the new data frame
D <- MDR_Dengue_map_pre %>%
  dplyr::select(all_of(selected_columns))
D1<- st_drop_geometry(D)
#write.csv(D1,'Weather_Data_with_New_Boundaries_2004_2023.csv')
#write.csv(filtered_data_summarized,'weather_station_6_variables.csv')

library(sf)
MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")
MDR_NEW$NAME_En<- toupper(MDR_NEW$NAME_En)
MDR_NEW$VARNAME<- toupper(MDR_NEW$VARNAME)



 
 matching_district <- (MDR_NEW$NAME_2 %in% filtered_data_summarized$NAME_2 )
  matching_district 

# Display the matching values
matched_names <- unique(MDR_NEW$NAME_2[matching_district])

matched_names

Non_matching<- unique(MDR_NEW$NAME_2[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)





MDR_Dengue_map_weather <- left_join(MDR_NEW, filtered_data_summarized_without_2023_year, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))

dim(MDR_Dengue_map_weather)

Dengue_cases<- MDR_Dengue_map_Final

MDR_Dengue_map_weather$NAME_En<- toupper(MDR_Dengue_map_weather$NAME_En)
MDR_Dengue_map_weather$VARNAME<- toupper(MDR_Dengue_map_weather$VARNAME)



Final_set<- inner_join(socio_dengue_pop, MDR_Dengue_map_weather, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="Year","month"="Month","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))

dim(Final_set)

# Assuming your existing data frame is named existing_data_frame
# Assuming you want to select specific columns based on the provided names

# Columns to select
selected_columns <- c(
   "province", "district", "year", "month", "Dengue", "SevereDHF",
   "NAME_1", "NAME_2", "TYPE_2",
  "ENGTYPE",   "t2m_avg",
  "t2m_max", "t2m_min", "tp_accum","rh_avg",   "rh_min" ,  "ws_avg",
  "total_population" ,  "Population_density"  ,              "Outmigration_Rate"  ,             
 "Inmigration_Rate"   ,              "NetImmigration_Rate"  ,            "Poverty_Rate",                    
 "Hygienic_Water_Access",            "Monthly_Average_Income_Percapita", "Total_Passenger",                 
 "Hygienic_Toilet_Access",           "Urbanization_Rate" 
)

# Create the new data frame
socio_dengue_pop_weather_station <- Final_set %>%
  dplyr::select(all_of(selected_columns))


###Save the results form the different data sets in one file 
```

##Preventive factors

```{r setup, include=FALSE}
file_path <- "C:\\Users\\uqwareed\\OneDrive - The University of Queensland\\Re District-level data_DUNG\\Preventive_factor_dengue_cases.xlsx"

Ento <- read_excel(file_path, sheet = 2)
Ento<- Ento[Ento$year >= 2004 & Ento$year < 2023, -c(3, 4, 5)]

Labo <- read_excel(file_path, sheet = 3)
Labo<- Labo[Labo$year >= 2004 & Labo$year < 2023, ]

Intervention<- read_excel(file_path, sheet = 4,na = "NA")
Intervention<- Intervention[Intervention$year>=2004 & Intervention$year < 2023,-c(3,4,5)]


# Capitalize the first letter of each word in District and Province columns
Ento$District <- str_to_title(Ento$District)
Ento$Province <- str_to_title(Ento$Province)
Ento$prov_dis <- paste0(Ento$District, " ", Ento$Province)

# Repeat the same for Labo data frame
Labo$District <- str_to_title(Labo$District)
Labo$Province <- str_to_title(Labo$Province)
Labo$prov_dis <- paste0(Labo$District, " ", Labo$Province)

# Repeat the same for Intervention data frame
Intervention$District <- str_to_title(Intervention$District)
Intervention$Province <- str_to_title(Intervention$Province)
Intervention$prov_dis <- paste0(Intervention$District, " ", Intervention$Province)


Labo_data_summarized<- Labo %>%  group_by( year,Month, District,Province)%>%
  dplyr::mutate(
    District = case_when(
      District %in% c("Mo Cay Nam", "Mo Cay Bac") & Province == "Ben Tre" ~ "Mo Cay",
      District %in% c("Co Do", "Thoi Lai",'Can Tho City') & Province == "Can Tho" ~ "Co Do",
      District %in% c("Hong Ngu City", "Hong Ngu Town") & Province == "Dong Thap" ~ "Hong Ngu",
      District %in% c("Nga Bay City", "Nga Bay Town", "Phung Hiep") & Province == "Hau Giang" ~ "Nga Bay",
      District %in% c( "Long My" ,"Long My City", "Long My Town") & Province == "Hau Giang" ~ "Long My",
      District %in% c("An Minh", "Vinh Thuan", "U Minh Thuong", "An Bien") & Province == "Kien Giang" ~ "An Minh",
      District %in% c("Kien Luong","Kien Tuong", "Giang Thanh") & Province == "Kien Giang" ~ "Kien Luong",
      District %in% c("Moc Hoa" , "Kien Tuong","Thi Xa Kien Tuong") & Province == "Long An" ~ "Moc Hoa",
      District %in% c("Long Phu", "My Xuyen", "Tran De") & Province == "Soc Trang" ~ "Long Phu",
      District %in% c("Chau Thanh", "My Tu") & Province == "Soc Trang" ~ "My Tu",
      District %in% c("Cai Lay", "Cai Lay City" ,"Tx. Cai Lay")  & Province == "Tien Giang" ~ "Cai Lay",
      District %in% c("Go Cong Dong", "Go Cong Tay", "Tan Phu Dong") & Province == "Tien Giang" ~ "Go Cong Dong",
      District %in% c("Thi Xa Duyen Hai","Duyen Hai","Duyen Hai City", "Duyen Hai Town","Tra Cu" ) & Province == "Tra Vinh" ~ "Tra Cu",
      District %in% c("Binh Minh Town", "Binh Tan") & Province == "Vinh Long" ~ "Binh Minh",
      District %in% c("Vinh Loi", "Hoa Binh") & Province == "Bac Lieu" ~ "Vinh Loi",
      TRUE ~ District
    )
  ) %>%
  summarise(
  `Total test` = sum(`Total test`, na.rm = TRUE),
  `No. DEN1` = sum(`No. DEN1`, na.rm = TRUE),
  `No. DEN2` = sum(`No. DEN2`, na.rm = TRUE),
  `No. DEN3` = sum(`No. DEN3`, na.rm = TRUE),
  `No. DEN4` = sum(`No. DEN4`, na.rm = TRUE)
) %>%
  ungroup()
# Create a new data frame to store the interpolated data

dim(Labo_data_summarized)

#repeated the same value for each month 

Labo_data <- Labo_data_summarized %>%
  arrange(year, Month)

#Join it with the new shape file
Labo_data$District<- toupper(Labo_data$District)
Labo_data$Province<- toupper(Labo_data$Province)

    MDR_NEW$VARNAME = toupper(MDR_NEW$VARNAME)
     MDR_NEW$NAME_En = toupper( MDR_NEW$NAME_En)
  

MDR_NEW <- MDR_NEW %>%
  mutate(NAME_En = gsub("CAM", "CA", NAME_En))
        

setdiff(MDR_NEW$NAME_En,Labo_data$Province)

setdiff(MDR_NEW$VARNAME,toupper(Labo_data$District))

setdiff(toupper(Labo_data$District),MDR_NEW$VARNAME)
# Create a mapping between old names and new names
district_mapping <- c(  "TP. BEN TRE" ="BEN TRE",
                      "SOC TRANG" = "SOC TRANG CITY" ,
                      "RACH GIA"= "RACH GIA CITY",
                     "SOC TRANG" = "SOC TRANG CITY" ,
                      "TRA VINH" = "TRA VINH CITY" ,
                     "CAO LANH" = "CAO LANH DISTRICT",
                     "TP MY THO"  ="MY THO",
                     "TP VINH LONG"="VINH LONG CITY",
                     "TRA VINH"="TRA VINH CITY",
                     "CAN THO CITY"="CAN THO",
                     "VI THANH CITY"="VI THANH"
                     )

 
# Replace the names in filled_data using the mapping
Labo_data <- Labo_data%>%
  dplyr::mutate(
   District  = ifelse(toupper(District)  %in% names(district_mapping), district_mapping[District ], District )
  )

Labo_data <- Labo_data %>%
  mutate(District = if_else(District == "TP. BEN TRE", "BEN TRE", District))

setdiff(toupper(MDR_NEW$VARNAME),toupper(Labo_data$District))
setdiff(toupper(Labo_data$District),toupper(MDR_NEW$VARNAME))

Labo_data$Province_District<- paste0(Labo_data$Province,"_",Labo_data$District)
MDR_NEW$province_district<- toupper(paste(MDR_NEW$NAME_En, MDR_NEW$VARNAME, sep = "_"))

setdiff(MDR_NEW$province_district,Labo_data$Province_District)
setdiff(Labo_data$Province_District,MDR_NEW$province_district)


 matching_district <- (MDR_NEW$province_district %in% Labo_data$Province_District )
  matching_district

   matching_district <- (Labo_data$Province_District %in% MDR_NEW$province_district  )
  matching_district
# Display the matching values
matched_names <- unique(Labo_data$Province_District[matching_district])

matched_names

Non_matching<- unique(Labo_data$Province_District[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)


MDR_Dengue_Labo <- left_join(MDR_NEW, Labo_data
                             , by = c("VARNAME"="District","NAME_En"="Province","province_district"="Province_District"))
#MDR_Dengue_pop$total_population<- ceiling(MDR_Dengue_pop$total_population)
dim(MDR_Dengue_Labo)
##The Labo data is not clloected fro each months so its dimension did not match the dengue cases.
MDR_Dengue_Labo <- sf::st_drop_geometry( MDR_Dengue_Labo )%>%
  distinct(year, Month, province_district, .keep_all = TRUE)

dim(MDR_Dengue_Labo)


MDR_Dengue_Labo$NAME_En<- toupper(MDR_Dengue_Labo$NAME_En)
MDR_Dengue_Labo$VARNAME<- toupper(MDR_Dengue_Labo$VARNAME)

Dengue_cases<- MDR_Dengue_map_Final %>%
  mutate(province = gsub("CAM", "CA", province))

Final_set_labo<- left_join(Dengue_cases, MDR_Dengue_Labo, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="year","month"="Month"))

dim(Final_set_labo)

##Inversigate the NA 

selected_columns <- c("province" ,         "district" ,"year","month", "Dengue","SevereDHF" ,
  "Total test"    ,    "No. DEN1"  ,        "No. DEN2"  ,        "No. DEN3"   ,       "No. DEN4"  
)

LABO <- Final_set_labo %>%
  dplyr::select(all_of(selected_columns))
##Labo is not conducted every month, so to match with data it assumed to be zeros 

LABO <- LABO %>%
  mutate_at(vars(`Total test`, `No. DEN1`, `No. DEN2`, `No. DEN3`, `No. DEN4`), ~ replace_na(., 0))

na_rows <- LABO[!complete.cases(LABO), ]

# Display the rows with NA values, if any
if (nrow(na_rows) > 0) {
  print("Rows with NA values:")
  print(na_rows)
} else {
  print("No rows with NA values found.")
}

#Ento data set
##Firstly fill all the months in the data 
 dim(Ento)

names(Ento)

library(mice)

Ento$HI_larvae<- as.numeric(Ento$HI_larvae)
Ento$BI_larvae<- as.numeric(Ento$BI_larvae)
Ento$CI_larvae<- as.numeric(Ento$CI_larvae)
Ento$DI<- as.numeric(Ento$DI)
Ento$HI_Mos<- as.numeric(Ento$HI_Mos)
# Specify the predictors and target variable
predictors <- setdiff(names(Ento), c("BI_larvae", "year", "Month", "Province", "District", "prov_dis"))
target <- "BI_larvae"
# Remove problematic variables from predictors if necessary
#predictors <- setdiff(predictors, c("communication_and/or_training"))

# Perform multiple imputation
imputed_data <- mice(Ento[, c(predictors, target)], method = "pmm", m = 5)

# Complete the imputation
completed_data1 <- complete(imputed_data)

dim(completed_data1)


# Fill NA values using ifelse
Ento$BI_larvae <- ifelse(
 is.na(Ento$BI_larvae),  # Condition: NA values
 ( completed_data1$BI_larvae),  # Value if NA
 Ento$BI_larvae  # Value if not NA
)


# Specify the predictors and target variable
predictors <- setdiff(names(Ento), c("CI_larvae", "year", "Month", "Province", "District", "prov_dis"))
target <- "CI_larvae"
# Remove problematic variables from predictors if necessary
#predictors <- setdiff(predictors, c("communication_and/or_training"))

# Perform multiple imputation
imputed_data <- mice(Ento[, c(predictors, target)], method = "pmm", m = 5)

# Complete the imputation
completed_data1 <- complete(imputed_data)

dim(completed_data1)
Ento$CI_larvae <- ifelse(
 is.na(Ento$CI_larvae),  # Condition: NA values
 ( completed_data1$CI_larvae),  # Value if NA
 Ento$CI_larvae  # Value if not NA
)


predictors <- setdiff(names(Ento), c("HI_larvae", "year", "Month", "Province", "District", "prov_dis"))
target <- "HI_larvae"
# Remove problematic variables from predictors if necessary
#predictors <- setdiff(predictors, c("communication_and/or_training"))

# Perform multiple imputation
imputed_data <- mice(Ento[, c(predictors, target)], method = "pmm", m = 5)

# Complete the imputation
completed_data1 <- complete(imputed_data)

dim(completed_data1)
Ento$HI_larvae <- ifelse(
 is.na(Ento$HI_larvae ),  # Condition: NA values
 ( completed_data1$HI_larvae ),  # Value if NA
 Ento$HI_larvae  # Value if not NA
)


predictors <- setdiff(names(Ento), c("DI", "year", "Month", "Province", "District", "prov_dis"))
target <- "DI"
# Remove problematic variables from predictors if necessary
#predictors <- setdiff(predictors, c("communication_and/or_training"))

# Perform multiple imputation
imputed_data <- mice(Ento[, c(predictors, target)], method = "pmm", m = 5)

# Complete the imputation
completed_data1 <- complete(imputed_data)

Ento$DI <- ifelse(
 is.na(Ento$DI ),  # Condition: NA values
 ( completed_data1$DI ),  # Value if NA
 Ento$DI  # Value if not NA
)


predictors <- setdiff(names(Ento), c("HI_Mos", "year", "Month", "Province", "District", "prov_dis"))
target <- "HI_Mos"
imputed_data <- mice(Ento[, c(predictors, target)], method = "pmm", m = 5)

completed_data1 <- complete(imputed_data)
Ento$HI_Mos <- ifelse(
 is.na(Ento$HI_Mos ),  
 ( completed_data1$HI_Mos ), 
 Ento$HI_Mos  
)

Ento_data_summarized <- Ento %>%  group_by( year,Month, District,Province)%>%
  mutate(
        District = 
      ifelse(
        (  District %in% c("Mo Cay Nam", "Mo Cay Bac")) &  Province %in% 'Ben Tre',
       "Mo Cay",
        ifelse(
          (  District %in%c("Co Do","Thoi Lai",'Can Tho City')) &  Province %in% 'Can Tho',
         "Co Do",
          ifelse(
            (  District %in% c("Hong Ngu City", "Hong Ngu Town"))  &  Province %in% 'Dong Thap',
           "Hong Ngu",
            ifelse(
              (  District %in% c("Nga Bay City", "Nga Bay Town","Phung Hiep","Nga Bay")) &  Province %in% c("Hau Giang",'Can Tho'),
              "Nga Bay",
              ifelse(
                (  District %in% c("Thi Xa Long My", "Long My" ,"Long My City", "Long My Town")) &  Province %in% "Hau Giang" ,
                "Long My",ifelse(
                  (District %in% c(  "Thi Xa Long My"  ,"Long My" ,"Long My City", "Long My Town")) &  Province %in% "Can Tho" ,
                "Long My",
                ifelse(
                  (  District %in%  c("An Minh", "Vinh Thuan",  "U Minh Thuong" ,"An Bien"  )) &  Province%in% "Kien Giang",
                  "An Minh",
                  ifelse(
                    (  District %in% c("Kien Luong", "Giang Thanh")) &  Province%in% "Kien Giang",
                   "Kien Luong",
                    ifelse(
                      (  District %in% c("Moc Hoa" , "Kien Tuong")) &  Province %in% "Long An",
                      "Moc Hoa" ,
                      ifelse(
                        (  District %in% c("Long Phu"  , "My Xuyen", "Tran De")) &  Province %in% "Soc Trang",
                        "Long Phu"  ,
                        ifelse(
                        (  District %in%c("Chau Thanh", "My Tu" )) &  Province %in% "Soc Trang",
                        "My Tu",
                        ifelse(
                          (  District %in% c("Cai Lay", "Cai Lay City" ,"Tx. Cai Lay") )&  Province %in% "Tien Giang",
                          "Cai Lay",
                          ifelse(
                            (  District %in% c("Go Cong Dong", "Go Cong Tay", "Tan Phu Dong" )) &  Province %in% "Tien Giang" ,
                            "Go Cong Dong",
                            ifelse(
                              (  District %in% c("Duyen Hai","Duyen Hai City", "Duyen Hai Town","Tra Cu" )) &  Province %in% "Tra Vinh" ,
                              "Tra Cu" ,
                              ifelse(
                                (  District %in%  c("Binh Minh Town", "Binh Tan" ) ) &  Province %in% "Vinh Long",
                               "Binh Minh",
                               ifelse(
       District %in% c("Vinh Loi", "Hoa Binh") &  Province %in% "Bac Lieu",
      "Vinh Loi",   District
                              )
      

                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
 ) ) )%>%
 
 summarise(
  BI_larvae = sum(as.numeric(BI_larvae), na.rm = TRUE),
  CI_larvae = sum(as.numeric(CI_larvae), na.rm = TRUE),
 HI_larvae = sum(as.numeric(HI_larvae), na.rm = TRUE),
 DI = sum(as.numeric(DI), na.rm = TRUE),
  HI_Mos = sum(as.numeric(HI_Mos), na.rm = TRUE),
) %>%
  ungroup()
# Create a new data frame to store the interpolated data
dim(Ento_data_summarized)

#repeated the same value for each month 

Ento_data <- Ento_data_summarized %>%
  arrange(year, Month)

Ento_data<- Ento_data[Ento_data$District != "NA"  , ]



#Join it with the new shape file

MDR_NEW <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")

MDR_NEW<- MDR_NEW %>%
  mutate(NAME_En = gsub("CAM", "CA", NAME_En))


MDR_NEW$VARNAME<- toupper(MDR_NEW$VARNAME)
Ento_data$District<- toupper(Ento_data$District)

MDR_NEW$NAME_En<- toupper(MDR_NEW$NAME_En)
Ento_data$Province<- toupper(Ento_data$Province)

setdiff(MDR_NEW$VARNAME,Ento_data$District)
setdiff(Ento_data$District,MDR_NEW$VARNAME)

# Create a mapping between old names and new names
district_mapping <- c(  "TP. BEN TRE" ="BEN TRE",
                      "SOC TRANG" = "SOC TRANG CITY" ,
                      "RACH GIA"= "RACH GIA CITY",
                     "SOC TRANG" = "SOC TRANG CITY" ,
                      "TRA VINH" = "TRA VINH CITY" ,
                     "CAO LANH" = "CAO LANH DISTRICT",
                     "TP MY THO"  ="MY THO",
                     "TP VINH LONG"="VINH LONG CITY",
                     "TRA VINH"="TRA VINH CITY",
                     "CAN THO CITY"="CAN THO",
                     "VI THANH CITY"="VI THANH",
                     "VINH LONG"="VINH LONG CITY",
                     "RACH GIA"="RACH GIA CITY"
                     )

# Replace the names in filled_data using the mapping
Ento_data <- Ento_data%>%
  mutate(
   District  = ifelse(District  %in% names(district_mapping), district_mapping[District ], District )
  )


setdiff(MDR_NEW$VARNAME,Ento_data$District)
setdiff(Ento_data$District,MDR_NEW$VARNAME)

##The Ento data set has a lot of missing districts 

MDR_Dengue_Ento <- left_join(MDR_NEW, Ento_data,
    by = c("VARNAME"="District","NAME_En"="Province"))

dim(MDR_Dengue_Ento)

not_matched <- anti_join(MDR_NEW, Ento_data,
    by = c("VARNAME"="District","NAME_En"="Province"))

# View the dimensions of the resulting data frame
dim(not_matched)

#Match it with dengue cases to remove any duplication.

Dengue_cases<- MDR_Dengue_map_Final %>%
  mutate(province = gsub("CAM", "CA", province))

Dengue_cases<- Dengue_cases %>%
  mutate(district = gsub("CAM", "CA", district))

MDR_Dengue_Ento$NAME_En<- toupper(MDR_Dengue_Ento$NAME_En)
MDR_Dengue_Ento$VARNAME<- toupper(MDR_Dengue_Ento$VARNAME)

MDR_Dengue_Ento<- MDR_Dengue_Ento%>%
  mutate(NAME_En = gsub("CAM", "CA", NAME_En))

MDR_Dengue_Ento<- MDR_Dengue_Ento %>%
  mutate(VARNAME = gsub("CAM", "CA", VARNAME))

Dengue_cases$province_district<- toupper(paste(Dengue_cases$province, Dengue_cases$district, sep = "_"))

setdiff(MDR_Dengue_Ento$province_district,Dengue_cases$province_district)

Final_set_Ento<- left_join(Dengue_cases, MDR_Dengue_Ento, by = c("year"="year","month"="Month","province"="NAME_En","district"="VARNAME"))

dim(Final_set_Ento)

selected_columns <- c(
   "province",          "district",          "year"   ,           "month"     ,        "Dengue" ,           "SevereDHF"  ,     "province", "Dengue","SevereDHF" ,          "BI_larvae",           "CI_larvae" ,         
 "HI_larvae" ,          "DI"  
)

ENTO <- Final_set_Ento %>%
  dplyr::select(all_of(selected_columns))



#For intervention data

Intervention<- read_excel(file_path, sheet = 4,na = "NA")

Intervention<- Intervention[Intervention$year>=2004 & Intervention$year < 2023,-c(3,4,5)]

Intervention$District <- str_to_title(Intervention$District)
Intervention$Province <- str_to_title(Intervention$Province)
Intervention$prov_dis <- paste0(Intervention$District, " ", Intervention$Province)

unique_combinations <- unique(Labo[, c("Province", "District", "prov_dis")])

# Create a data frame with all combinations of year and month
year_month <- expand.grid(
  year = unique(Intervention$year),
  month = 1:12
)

# Create a template data frame by merging year_month with unique_combinations
template_df <- merge(year_month, unique_combinations, by = NULL)

# Left join your data onto the template data frame
Intervention<- merge(template_df, Intervention, by = c("year", "month", "Province", "District", "prov_dis"), all.x = TRUE)

###Expand the data to inclde also the years from 2004 to 2014
library(dplyr)

# Create a data frame with all combinations of years, months, provinces, and districts
all_combinations <- expand.grid(year = 2004:2022,
                                month = 1:12
                                )

template_df <- merge(all_combinations, unique_combinations, by = NULL)

# Left join your data onto the template data frame
Intervention<- merge(template_df, Intervention, by = c("year", "month", "Province", "District", "prov_dis"), all.x = TRUE)

library(dplyr)

# Load the mice package for multiple imputation
library(mice)

# Specify the predictors and target variable
predictors <- setdiff(names(Intervention), c("breeding_site_elimination_campaign", "year", "month", "Province", "District", "prov_dis"))
target <- "breeding_site_elimination_campaign"
# Remove problematic variables from predictors if necessary
#predictors <- setdiff(predictors, c("communication_and/or_training"))

# Perform multiple imputation
imputed_data <- mice(Intervention[, c(predictors, target)], method = "pmm", m = 5)

# Complete the imputation
completed_data1 <- complete(imputed_data)

# Check if there are still missing values
sum(is.na(completed_data1$large_scale_spraying_for_epidemic_response))

dim(completed_data1)

# Create a logical condition to identify NA values in breeding_site_elimination_campaign
na_condition <- is.na(Intervention$breeding_site_elimination_campaign)

# Fill NA values using ifelse
Intervention$breeding_site_elimination_campaign <- ifelse(
 is.na(Intervention$breeding_site_elimination_campaign),  # Condition: NA values
  completed_data1$breeding_site_elimination_campaign,  # Value if NA
  Intervention$breeding_site_elimination_campaign  # Value if not NA
)

Intervention$active_spraying <- ifelse(
 is.na(Intervention$active_spraying),  # Condition: NA values
  completed_data1$active_spraying,  # Value if NA
  Intervention$active_spraying  # Value if not NA
)

Intervention$large_scale_spraying_for_epidemic_response <- ifelse(
 is.na(Intervention$large_scale_spraying_for_epidemic_response),  # Condition: NA values
  completed_data1$large_scale_spraying_for_epidemic_response,  # Value if NA
  Intervention$large_scale_spraying_for_epidemic_response  # Value if not NA
)

Intervention$communication_or_training <- ifelse(
 is.na(Intervention$communication_or_training),  # Condition: NA values
  completed_data1$communication_or_training,  # Value if NA
  Intervention$communication_or_training  # Value if not NA
)

Intervention$number_of_outbreak_detection <- ifelse(
 is.na(Intervention$number_of_outbreak_detection),  # Condition: NA values
  completed_data1$number_of_outbreak_detection,  # Value if NA
  Intervention$number_of_outbreak_detection  # Value if not NA
)

Intervention$number_of_outbreak_response <- ifelse(
 is.na(Intervention$number_of_outbreak_response),  # Condition: NA values
  completed_data1$number_of_outbreak_response,  # Value if NA
  Intervention$number_of_outbreak_response  # Value if not NA
)

Intervention_data_summarized <- Intervention %>%  group_by( year,month, District,Province)%>%
  mutate(
        District = 
      ifelse(
        (  District %in% c("Mo Cay Nam", "Mo Cay Bac")) &  Province %in% 'Ben Tre',
       "Mo Cay",
        ifelse(
          (  District %in%c("Co Do","Thoi Lai",'Can Tho City')) &  Province %in% 'Can Tho',
         "Co Do",
          ifelse(
            (  District %in% c("Hong Ngu City", "Hong Ngu Town","Thi Xa Hong Ngu"))  &  Province %in% 'Dong Thap',
           "Hong Ngu",
            ifelse(
              (  District %in% c("Nga Bay City", "Nga Bay Town","Phung Hiep","Nga Bay")) &  Province %in% c("Hau Giang",'Can Tho'),
              "Nga Bay",
              ifelse(
                (  District %in% c( "Thi Xa Long My","Long My" ,"Long My City", "Long My Town")) &  Province %in% "Hau Giang" ,
                "Long My",ifelse(
                  (District %in% c( "Long My" ,"Long My City", "Long My Town","Thi Xa Long My")) &  Province %in% "Can Tho" ,
                "Long My",
                ifelse(
                  (  District %in%  c("An Minh", "Vinh Thuan",  "U Minh Thuong" ,"An Bien"  )) &  Province%in% "Kien Giang",
                  "An Minh",
                  ifelse(
                    (  District %in% c("Kien Luong", "Giang Thanh")) &  Province%in% "Kien Giang",
                   "Kien Luong",
                    ifelse(
                      (  District %in% c("Moc Hoa" , "Kien Tuong","Thi Xa Kien Tuong")) &  Province %in% "Long An",
                      "Moc Hoa" ,
                      ifelse(
                        (  District %in% c("Long Phu"  , "My Xuyen", "Tran De")) &  Province %in% "Soc Trang",
                        "Long Phu"  ,
                        ifelse(
                        (  District %in%c("Chau Thanh", "My Tu" )) &  Province %in% "Soc Trang",
                        "My Tu",
                        ifelse(
                          (  District %in% c("Thi Xa Cai Lay","Cai Lay", "Cai Lay City" ,"Tx. Cai Lay") )&  Province %in% "Tien Giang",
                          "Cai Lay",
                          ifelse(
                            (  District %in% c("Go Cong Dong", "Go Cong Tay", "Tan Phu Dong" )) &  Province %in% "Tien Giang" ,
                            "Go Cong Dong",
                            ifelse(
                              (  District %in% c("Thi Xa Duyen Hai","Duyen Hai","Duyen Hai City", "Duyen Hai Town","Tra Cu" )) &  Province %in% "Tra Vinh" ,
                              "Tra Cu" ,
                              ifelse(
                                (  District %in%  c("Binh Minh Town", "Binh Tan" ) ) &  Province %in% "Vinh Long",
                               "Binh Minh",
                               ifelse(
       District %in% c("Vinh Loi", "Hoa Binh") &  Province %in% "Bac Lieu",
      "Vinh Loi",   District
                              )
      

                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
 ) ) )%>%
 
 summarise(
 breeding_site_elimination_campaign = sum(as.numeric(breeding_site_elimination_campaign), na.rm = TRUE),
 active_spraying = sum(as.numeric(active_spraying), na.rm = TRUE),
large_scale_spraying_for_epidemic_response = sum(as.numeric(large_scale_spraying_for_epidemic_response), na.rm = TRUE),
`communication_or_training`= sum(as.numeric(`communication_or_training`), na.rm = TRUE),
  number_of_outbreak_detection = sum(as.numeric(number_of_outbreak_detection), na.rm = TRUE),
number_of_outbreak_response = sum(as.numeric(number_of_outbreak_response), na.rm = TRUE)
) %>%
  ungroup()

dim(Intervention_data_summarized)
# Create a new data frame to store the interpolated data
#Ento_data_summarized<- na.omit(Ento_data_summarized)


Intervention_data <- Intervention_data_summarized %>%
  arrange(year, month)


Intervention_data<-Intervention_data[Intervention_data$District != "NA"  , ]

MDR_NEW <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Re District-level data_DUNG/MDR_NEW_Boundaries.shp")

Intervention_data$Province  <- ifelse(
  is.na(Intervention_data$Province) &Intervention_data$District == "CAO LANH",  # Condition: NA in Province and District is "CAO
  "DONG THAP",  # Value if condition is met
  Intervention_data$Province # Original value if condition is not met
)

MDR_NEW$VARNAME<- toupper(MDR_NEW$VARNAME)
Intervention_data$District<- toupper(Intervention_data$District)

MDR_NEW$NAME_En<- toupper(MDR_NEW$NAME_En)


MDR_NEW<- MDR_NEW %>%
  mutate(NAME_En = gsub("CAM", "CA", NAME_En))

Intervention_data$Province<- toupper(Intervention_data$Province)

setdiff(MDR_NEW$VARNAME,Intervention_data$District)
setdiff(Intervention_data$District,MDR_NEW$VARNAME)
# Create a mapping between old names and new names
district_mapping <- c(  "TP. BEN TRE" ="BEN TRE",
                      "SOC TRANG" = "SOC TRANG CITY" ,
                      "RACH GIA"= "RACH GIA CITY",
                    "THANH PHO SOC TRANG" = "SOC TRANG CITY" ,
                      "TRA VINH" = "TRA VINH CITY" ,
                     "CAO LANH" = "CAO LANH DISTRICT",
                     "TP MY THO"  ="MY THO",
                     "TP VINH LONG"="VINH LONG CITY",
                     "TRA VINH"="TRA VINH CITY",
                     "CAN THO CITY"="CAN THO",
                     "VI THANH CITY"="VI THANH",
                     "VINH LONG"="VINH LONG CITY",
                      "THANH PHO RACH GIA"="RACH GIA CITY",
                    "THI XA VI THANH"   = "VI THANH" ,
                    "THANH PHO CA MAU"= "CA MAU" ,
                    "THANH PHO TAN AN"  ="TAN AN",
                   "THANH PHO CAO LANH"= "CAO LANH DISTRICT",
                   "THANH PHO TRA VINH"="TRA VINH CITY" ,
                   "SA ĐEC"="SA DEC",
                   "THI XA GO CONG" ="GO CONG" ,
                   "THANH PHO VINH LONG"="VINH LONG CITY",
                   "THANH PHO BAC LIEU"="BAC LIEU" ,
                   "THANH PHO BEN TRE" = "BEN TRE" ,
                   "THANH PHO VINH LONG"="VINH LONG CITY"
                    )

# Replace the names in filled_data using the mapping
Intervention_data <- Intervention_data%>%
  mutate(
   District  = ifelse(District  %in% names(district_mapping), district_mapping[District ], District )
  )



setdiff(MDR_NEW$VARNAME,Intervention_data$District)
setdiff(Intervention_data$District,MDR_NEW$VARNAME)

# Replace NA values with a similar name
#similar_name <- "CAO LANH CITY"
#Intervention_data$District[is.na(Intervention_data$District)] <- similar_name
# Fill NA values in Province column based on District
setdiff(Intervention_data$Province,MDR_NEW$NAME_En)



Intervention_data$Province_District<- paste0(Intervention_data$Province,"_",Intervention_data$District)
MDR_NEW$province_district<- toupper(paste(MDR_NEW$NAME_En, MDR_NEW$VARNAME, sep = "_"))

setdiff(MDR_NEW$province_district,Intervention_data$Province_District)
setdiff(Intervention_data$Province_District,MDR_NEW$province_district)

# Assuming 'expanded_data' is your tibble
MDR_NEW<- MDR_NEW %>%
  mutate(VARNAME = gsub("CAM", "CA", VARNAME))

MDR_NEW<- MDR_NEW %>%
  mutate(province_district = gsub("CAM", "CA", province_district))



 matching_district <- (MDR_NEW$province_district %in% Intervention_data$Province_District )
  matching_district

   matching_district <- (Intervention_data$Province_District %in% MDR_NEW$province_district  )
  matching_district
# Display the matching values
matched_names <- unique(Intervention_data$Province_District[matching_district])

matched_names

Non_matching<- unique(Intervention_data$Province_District[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)

MDR_Dengue_Intervention <- left_join(MDR_NEW, Intervention_data
                             , by = c("VARNAME"="District","NAME_En"="Province","province_district"="Province_District"))
#MDR_Dengue_pop$total_population<- ceiling(MDR_Dengue_pop$total_population)
dim(MDR_Dengue_Intervention)
##The Intervention data is not collected fro each months so its dimension did not match the dengue cases.


MDR_Dengue_Intervention$NAME_En<- toupper(MDR_Dengue_Intervention$NAME_En)
MDR_Dengue_Intervention$VARNAME<- toupper(MDR_Dengue_Intervention$VARNAME)

Dengue_cases<- Dengue_cases %>%
  mutate(province = gsub("CAM", "CA", province))

Final_set<- inner_join(Dengue_cases, MDR_Dengue_Intervention, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="year","month"="month"))

dim(Final_set)
Final_set <- distinct(Final_set)

# Check dimensions after removing duplicates
dim(Final_set)
Final_set <- sf::st_drop_geometry( Final_set )%>%
  distinct(year, month, province_district, .keep_all = TRUE)

library(dplyr)

# Perform an anti-join to find rows that did not match
not_matched <- anti_join(Dengue_cases, MDR_Dengue_Intervention, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="year","month"="month"))

# View the dimensions of the resulting data frame
dim(not_matched)

##Inversigate the NA 

library(dplyr)

# Extract rows with NA values


selected_columns <- c(
   "NAME_1.x"   ,       "NAME_2.x", "TYPE_2.x","year","month",
   "ENGTYPE.x" ,"district",     "province", "Dengue","SevereDHF" ,
 "breeding_site_elimination_campaign"  ,       "active_spraying",                           
"large_scale_spraying_for_epidemic_response", "communication_or_training",                 
 "number_of_outbreak_detection"    ,           "number_of_outbreak_response" 
)

D3 <- Final_set %>%
  dplyr::select(all_of(selected_columns))



```