---
title: "Data_cleaning"
output: html_document
date: "2024-06-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This file includes the process to get the final data  the we use it for monthly analysis
The data cleaning process including obtain the data from different resources, here these data has been arranged to match our new boundaries
To see how we obtained the new boundaries please see shapefile_new_from_2004.rmd
The data process in  this file includes dengue cases, population, weather variable form ERA5, that found to be important for ensemble model 


##Dengue cases
```{r,warning=false }
library(stringr)
library(ggplot2)
library(dplyr)
library(readxl) 
library(dplyr)
library(readr)
library(readxl)
library(ggplot2)
library(spdep)
library(sp)
library(MASS)
library(surveillance)
library(tidyr)
library(splines)
library(tidyverse)
library(tsibble)
library(sf)
library(scales)
library(RColorBrewer)
library(viridis)
library(readxl)
Dengue<- read_excel("C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Raw_data/20231130_ED_MDR Dengue district data_2000-2023_original modification_updated 20May24.xlsx", sheet=1)
Dengue_filtered_data <- Dengue[Dengue$year >= 2004, ]
dim(Dengue_filtered_data)

unique(Dengue_filtered_data$district)

sum(is.na(Dengue_filtered_data$Dengue))
sum(is.na(Dengue_filtered_data$SevereDHF))


#Fill the missing values using two ways
##Use the average fro the same month in previous years.

Dengue_filtered_data <- Dengue_filtered_data %>%
  arrange(district, province, year, month)

# Define a function to fill missing values with the average from the same month in preceding years
fill_missing_with_average <- function(data, column) {
  data %>%
    group_by(district, province, month) %>%
    mutate({{column}} := ifelse(is.na({{column}}), mean({{column}}[year %in% (max(year) - 1):(min(year) + 1)], na.rm = TRUE), {{column}})) %>%
    ungroup()
}

# Apply the function to the columns "Dengue" and "SevereDHF"
Dengue_filtered_data <- Dengue_filtered_data %>%
  fill_missing_with_average(., Dengue) %>%
  fill_missing_with_average(., SevereDHF)

Dengue_filtered_data$Dengue<- ceiling(Dengue_filtered_data$Dengue)
Dengue_filtered_data$SevereDHF<- ceiling(Dengue_filtered_data$SevereDHF)
# Count the number of remaining missing values
missing_dengue <- sum(is.na(Dengue_filtered_data$Dengue))
missing_severe_dhf <- sum(is.na(Dengue_filtered_data$SevereDHF))

# Print the result
cat("Missing values in Dengue:", missing_dengue, "\n")
cat("Missing values in SevereDHF:", missing_severe_dhf, "\n")

#####Match the boundaries 

Dengue_filtered_data$district<- toupper(Dengue_filtered_data$district)
# Assuming your data frame is named filtered_data
Dengue_filtered_data_summarized <- Dengue_filtered_data %>%  group_by(province, district, year, month)%>%
  mutate(
    district = ifelse(
      (district %in% c("VINH LOI", "HOA BINH")) & (province == "BAC LIEU"),
      "VINH LOI",
      ifelse(
        (district %in% c("MO CAY","MO CAY NAM", "MO CAY BAC")) & (province == "BEN TRE"),
        "MO CAY",
        ifelse(
          (district %in% c("CO DO", "THOI LAI")) & (province == "CAN THO"),
          "CO DO",
          ifelse(
            (district %in% c("HONG NGU DISTRICT", "HONG NGU CITY" )) & (province ==  "DONG THAP" ),
            "HONG NGU",
            ifelse(
              (district %in% c( "NGA BAY CITY" , "PHUNG HIEP" )) & (province ==  "HAU GIANG" ),
              "NGA BAY",
              ifelse(
                (district %in% c( "LONG MY"  , "LONG MY TOWN" )) & (province ==  "HAU GIANG" ) ,
                "LONG MY",
                ifelse(
                  (district %in% c(  "AN MINH",  "U MINH THUONG"  , "AN BIEN" , "VINH THUAN"  )) & (province == "KIEN GIANG" ),
                  "AN MINH",
                  ifelse(
                    (district %in% c(  "KIEN LUONG",   "GIANG THANH"  )) & (province == "KIEN GIANG" ),
                    "KIEN LUONG",
                    ifelse(
                      (district %in% c(  "KIEN TUONG",   "MOC HOA"  )) & (province == "LONG AN" ),
                      "MOC HOA",
                      ifelse(
                        (district %in% c(  "LONG PHU" ,   "MY XUYEN","TRAN DE"   )) & (province == "SOC TRANG" ),
                        "LONG PHU",
                        ifelse(
                          (district %in% c(  "CHAU THANH" ,   "MY TU" )) & (province == "SOC TRANG" ),
                          "MY TU",
                          ifelse(
                            (district %in% c(  "CAI LAY" ,   "CAI LAY CITY" ,"CAI LAY DISTRICT"  )) & (province == "TIEN GIANG" ),
                            "CAI LAY",
                            ifelse(
                              (district %in% c(  "GO CONG TAY" ,   "GO CONG DONG", "TAN PHU DONG"   )) & (province == "TIEN GIANG" ),
                              "GO CONG DONG",
                              ifelse(
                                (district %in% c(  "DUYEN HAI DISTRICT" ,   "DUYEN HAI", "DUYEN HAI TOWN" ,"TRA CU"  )) & (province == "TRA VINH"   ),
                                "TRA CU",
                                ifelse(
                                  (district %in% c(    "BINH MINH", "BINH TAN"   )) & (province == "VINH LONG" ),
                                  "BINH MINH",
                                  district
                                )
                              )
                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    ) )%>%
  
  summarise(Dengue = sum(Dengue,na.rm = TRUE), SevereDHF = sum(SevereDHF,na.rm = TRUE)) %>%
  ungroup()

Dengue_filtered_data_summarized <- Dengue_filtered_data_summarized %>%
  mutate(
    district = ifelse(district ==  "RACH GIA CI TY"  , "RACH GIA CITY" , district)
  )



MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Data/shapefiles/MDR_NEW_Boundaries_Final.shp") 

MDR_NEW$VARNAME  <- toupper(MDR_NEW$VARNAME)
Dengue_filtered_data_summarized$district<- toupper(Dengue_filtered_data_summarized$district)
unique(MDR_NEW$VARNAME)

Dengue_filtered_data_summarized$province<- toupper(Dengue_filtered_data_summarized$province)
MDR_NEW$NAME_En<- toupper( MDR_NEW$NAME_En)

matching_district <- MDR_NEW$VARNAME %in% Dengue_filtered_data_summarized$district &
  MDR_NEW$NAME_En %in% Dengue_filtered_data_summarized$province

matching_district 
unique(Dengue_filtered_data_summarized$district%in% MDR_NEW$VARNAME )
# Identify values in filtered_data_summarized$district that are not in MDR_NEW$VARNAME
not_in_MDR_NEW <- setdiff(Dengue_filtered_data_summarized$district, MDR_NEW$VARNAME)

setdiff( MDR_NEW$VARNAME,Dengue_filtered_data_summarized$district)
setdiff(Dengue_filtered_data_summarized$district, MDR_NEW$VARNAME)
# Print or inspect the values
print(not_in_MDR_NEW)

# Display the matching values
matched_names <- unique(MDR_NEW$VARNAME[matching_district])

Non_matching<- unique(MDR_NEW$VARNAME[!matching_district])

Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)


# Display the indices or positions where matching_district is FALSE
q<- (unique(paste(MDR_NEW$VARNAME, MDR_NEW$NAME_En, sep = "_")))
#subset(MDR_NEW,MDR_NEW$VARNAME=="HONG NGU" )
#subset(MDR_2,MDR_2$VARNAME_2=="HONG NGU" )
non_unique_combinations <- MDR_NEW[duplicated(paste(MDR_NEW$VARNAME, MDR_NEW$NAME_En, sep = "_")) | duplicated(paste(MDR_NEW$VARNAME, MDR_NEW$NAME_En, sep = "_"), fromLast = TRUE), ]

MDR_Dengue_map_pre <- left_join( Dengue_filtered_data_summarized,MDR_NEW, by = c( "district"="VARNAME" ,"province"="NAME_En"))

dim(MDR_Dengue_map_pre)
names(MDR_Dengue_map_pre)
selected_columns <- c(
  "province",  "district",  "year" ,     "month"   ,  "Dengue",    "SevereDHF"
  ,  "NAME_1",       "NAME_2"   ,"TYPE_2"    ,"ENGTYPE" ,     "ID")

# Create the new data frame
MDR_Dengue_map_pre <- MDR_Dengue_map_pre %>%
  dplyr::select(all_of(selected_columns))

MDR_Dengue_map_Final<- st_drop_geometry(MDR_Dengue_map_pre)

BAC_LIEU_N <- MDR_Dengue_map_Final[MDR_Dengue_map_Final$district=='BAC LIEU',]

ggplot(BAC_LIEU_N, aes(x = month, y = Dengue, group = year, color = factor(year))) +
  geom_line() +
  labs(title = "Monthly Dengue Cases in Bac Lieu",
       x = "Month",
       y = "Number of Dengue Cases",
       color = "Year") +
  theme_minimal()


BAC_LIEU_N$date <- as.Date(paste(BAC_LIEU_N$year, BAC_LIEU_N$month, "1", sep = "-"))

# Create a time series plot
ggplot(BAC_LIEU_N, aes(x = date, y = Dengue)) +
  geom_line() +
  labs(title = "Time Series of Dengue Cases in Bac Lieu",
       x = "Date",
       y = "Number of Dengue Cases") +
  theme_minimal()
```



## Populaion data

```{r setup, include=FALSE}
file_path <- "C:\\Users\\uqwareed\\OneDrive - The University of Queensland\\Re District-level data_DUNG\\MDR_pop_district-Eng version- LT.xlsx"

df <- read_excel(file_path, sheet = 1)
df<- df[,c(3,5,6,8,9)]
df<- df[df$Year >=2004,]
df$District <- str_to_title(df$District)
df$Province<- str_to_title(df$Province)
df$prov_dis <- paste0(df$District,"", df$Province)

pop_filtered_data_summarized <- df %>%  group_by( Year, District,Province)%>%
  mutate(
        District = 
      ifelse(
        (  District %in% c("Mo Cay Nam", "Mo Cay Bac")) &  Province %in% 'Ben Tre',
       "Mo Cay",
        ifelse(
          (  District %in%c("Co Do","Thoi Lai")) &  Province %in% 'Can Tho',
         "Co Do",
          ifelse(
            (  District %in% c("Hong Ngu City", "Hong Ngu Town"))  &  Province %in% 'Dong Thap',
           "Hong Ngu",
            ifelse(
              (  District %in% c("Nga Bay City", "Nga Bay Town","Phung Hiep")) &  Province %in% "Hau Giang",
              "Nga Bay",
              ifelse(
                (  District %in% c("Long My", "Long My Town")) &  Province %in% "Hau Giang" ,
                "Long My",
                ifelse(
                  (  District %in%  c("An Minh", "Vinh Thuan",  "U Minh Thuong" ,"An Bien"  )) &  Province%in% "Kien Giang",
                  "An Minh",
                  ifelse(
                    (  District %in% c("Kien Luong", "Giang Thanh")) &  Province%in% "Kien Giang",
                   "Kien Luong",
                    ifelse(
                      (  District %in% c("Moc Hoa" , "Kien Tuong Town")) &  Province %in% "Long An",
                      "Moc Hoa" ,
                      ifelse(
                        (  District %in% c("Long Phu"  , "My Xuyen", "Tran De")) &  Province %in% "Soc Trang",
                        "Long Phu"  ,
                        ifelse(
                        (  District %in%c("Chau Thanh", "My Tu" )) &  Province %in% "Soc Trang",
                        "My Tu",
                        ifelse(
                          (  District %in% c("Cai Lay", "Cai Lay City" ) )&  Province %in% "Tien Giang",
                          "Cai Lay",
                          ifelse(
                            (  District %in% c("Go Cong Dong", "Go Cong Tay", "Tan Phu Dong" )) &  Province %in% "Tien Giang" ,
                            "Go Cong Dong",
                            ifelse(
                              (  District %in% c("Duyen Hai", "Duyen Hai Town","Tra Cu" )) &  Province %in% "Tra Vinh" ,
                              "Tra Cu" ,
                              ifelse(
                                (  District %in%  c("Binh Minh Town", "Binh Tan" ) ) &  Province %in% "Vinh Long",
                               "Binh Minh",
                               ifelse(
       District %in% c("Vinh Loi", "Hoa Binh") &  Province %in% "Bac Lieu",
      "Vinh Loi",   District
                              )
      

                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  ) )%>%
 
  summarise(
    total_population = sum(Total_population,na.rm = TRUE),Population_density=sum(Population_density,na.rm=TRUE))%>%
  ungroup()

pop_filtered_data_summarized$Province_District<-paste0(pop_filtered_data_summarized$Province,"_",pop_filtered_data_summarized$District)




# Create a new data frame to store the interpolated data
#repeated the same value for each month 


# Assuming filled_data is your original data frame
# Create a new data frame with the 'Month' column
expanded_data <- expand.grid(Month = 1:12, District = unique(pop_filtered_data_summarized$District), Province = unique(pop_filtered_data_summarized$Province))

# Merge with filled_data to get the required columns
expanded_data <- merge(expanded_data, pop_filtered_data_summarized, by = c("District", "Province"))

# Use tidyr's complete function to create all combinations of Year, Month, District, and Province
expanded_data <- arrange(complete(expanded_data, Year, Month, nesting(District, Province)))
expanded_data <- expanded_data %>%
  arrange(Year, Month)
# Print the expanded_data
print(expanded_data)

data_2021 <- expanded_data %>%
  filter(Year == 2021 & District == "Cao Lanh City" & Province == "Dong Thap") 

# Check the extracted 2021 data
print(data_2021)

# Fill in the missing 2022 data with the 2021 data
expanded_data <- expanded_data %>%
  mutate(total_population = ifelse(Year == 2022 & District == "Cao Lanh City" & Province == "Dong Thap",
                                   data_2021$total_population[match(Month, data_2021$Month)],
                                   total_population),
         Population_density = ifelse(Year == 2022 & District == "Cao Lanh City" & Province == "Dong Thap",
                                     data_2021$Population_density[match(Month, data_2021$Month)],
                                     Population_density))


na_rows <- expanded_data[is.na(expanded_data$total_population), ]

##fill the one year missing pop data for CAO LANH CITY in 2022 to be like in 2021

#Join it with the new shape file


MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Data/shapefiles/MDR_NEW_Boundaries_Final.shp") 

MDR_NEW$VARNAME<- toupper(MDR_NEW$VARNAME)
expanded_data$District<- toupper(expanded_data$District)

MDR_NEW$NAME_En<- toupper(MDR_NEW$NAME_En)
expanded_data$Province<- toupper(expanded_data$Province)

setdiff(MDR_NEW$VARNAME,expanded_data$District)
setdiff(expanded_data$District,MDR_NEW$VARNAME)
# Create a mapping between old names and new names
district_mapping <- c(  "CAO LANH" ="CAO LANH DISTRICT",
                      "GIA RAI TOWN" = "GIA RAI" ,
                      "RACH GIA"= "RACH GIA CITY",
                      "NGA NAM TOWN" ="NGA NAM" ,
                     "SOC TRANG" = "SOC TRANG CITY" ,
                      "TRA VINH" = "TRA VINH CITY" ,
                      "VINH LONG" = "VINH LONG CITY",
                     "TAN CHAU TOWN"  ="TAN CHAU")

# Replace the names in filled_data using the mapping
expanded_data <- expanded_data%>%
  mutate(
   District  = ifelse(District  %in% names(district_mapping), district_mapping[District ], District )
  )

##Fill the missing values 


setdiff(MDR_NEW$VARNAME,expanded_data$District)
setdiff(expanded_data$District,MDR_NEW$VARNAME)

expanded_data$Province_District<- paste0(expanded_data$Province,"_",expanded_data$District)
MDR_NEW$province_district<- toupper(paste(MDR_NEW$NAME_En, MDR_NEW$VARNAME, sep = "_"))

setdiff(MDR_NEW$province_district,expanded_data$Province_District)
setdiff(expanded_data$Province_District,MDR_NEW$province_district)

# Assuming 'expanded_data' is your tibble
MDR_NEW<- MDR_NEW %>%
  mutate(VARNAME = gsub("CAM", "CA", VARNAME))

MDR_NEW<- MDR_NEW %>%
  mutate(province_district = gsub("CAM", "CA", province_district))

# View the updated tibble
print(expanded_data)


 matching_district <- (MDR_NEW$province_district %in% expanded_data$Province_District )
  matching_district

# Display the matching values
matched_names <- unique(MDR_NEW$province_district[matching_district])

matched_names

Non_matching<- unique(MDR_NEW$province_district[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)


MDR_pop <- left_join(MDR_NEW, expanded_data, by = c("province_district"="Province_District"))

dim(MDR_pop)


selected_columns <- c(
   "NAME_1", "NAME_2", "TYPE_2","Year","Month","VARNAME",
  "ENGTYPE",  "total_population","District",     "Province","Population_density"
  
)

Dengue_pop<- inner_join(MDR_Dengue_map_Final, MDR_pop, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="Year","month"="Month","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))


dim(Dengue_pop)

# Assuming your existing data frame is named existing_data_frame
# Assuming you want to select specific columns based on the provided names

# Columns to select
selected_columns <- c(
   "province", "district", "year", "month", "Dengue", "SevereDHF",
   "NAME_1", "NAME_2", "TYPE_2",
  "ENGTYPE",   "total_population","Population_density"
)

# Create the new data frame
Final_subset_dengue_pop <- Dengue_pop %>%
  dplyr::select(all_of(selected_columns))

```



##Weather variables from ERA5
ws_avg: daily average wind speed (m/s)
t2m_avg: daily average temperature (in deg K = deg C + 273.15)
rh_avg: daily average relative humidity (%)
 ws_max: daily maximum wind speed (m/s)
 t2m_max: daily maximum temperature (in deg K = deg C + 273.15)
 rh_max: daily maximum relative humidity (%)
 ws_min: daily minimum wind speed (m/s)
t2m_min: daily minimum temperature (in deg K = deg C + 273.15)
rh_min: daily minimum relative humidity (%)
 tp_accum: accumulated total precipitation (m)
 

```{r,Warning=False}


##These data can be found in Van Github per request

directory_path <- "C:/Users/uqwareed/OneDrive - The University of Queensland/ERA5land_VN_districs-final/ERA5land_VN_districs-main/MekongDelta/metdata/ERA5/ERA5land_data"

# Get a list of all files in the directory with a .csv extension
files <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

##These lines will take long time to run 
# Initialize an empty data frame to store the results
result_df <- data.frame()
# Iterate through the list of files and read each CSV file
for (file in files) {
  # Extract year and month from the file name
  file_parts <- strsplit(basename(file), "\\.")[[1]]
  year_month <- substr(file_parts[1], 1, 6)
  year <- substr(year_month, 1, 4)
  month <- substr(year_month, 5, 6)

  # Read the CSV file
  df <- read_csv(file, col_types = cols(.default = "c"))
df1 <- apply(df[-c(1,2),2:ncol(df)], 2, function(x) as.numeric(as.character(x)))
  # Calculate the mean for each column
  means <- colMeans(df1, na.rm = TRUE)

  # Combine year, month, and mean into a data frame
  summary_data <- data.frame(
    Year = as.integer(year),
    Month = as.integer(month),ID= as.vector(as.character(df[1,2:ncol(df)])),
    Mean = means,names= as.vector(as.character(df[1,2:ncol(df)]))
  )

  # Append the summary data to the result data frame
  result_df <- bind_rows(result_df, summary_data)
}

# Print the resulting data frame
#print(result_df)
#saveRDS(result_df,'result_df.rds')


##Reshape the data set

result_df<- readRDS('C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Raw_data/result_df.rds')
# Assuming result_df is your data frame
unique_names <- unique(result_df$names)

library(dplyr)

new_df <- result_df %>%
  group_by(Year, Month) %>%
  summarize(
    ws_avg = Mean[names == "ws_avg"],
    ws_max = Mean[names == "ws_max"],
    ws_min = Mean[names == "ws_min"],
    t2m_avg = Mean[names == "t2m_avg"],
    t2m_max = Mean[names == "t2m_max"],
    t2m_min = Mean[names == "t2m_min"],
    rh_avg = Mean[names == "rh_avg"],
    rh_max = Mean[names == "rh_max"],
    rh_min = Mean[names == "rh_min"],
    tp_accum = Mean[names == "tp_accum"]
  ) %>%
  ungroup()  # Remove grouping for the new data frame

print(new_df)

#add the ID
result_df <- new_df %>%
  mutate(ID = rep(0:133, length.out = n()))

# View the updated data frame
print(result_df)
tail(result_df)


#Match the ID with the  VARNAME_2 name

VARNAME_2_inf<- read.csv('C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Raw_data/district_infomation.csv')

VARNAME_2_inf$X<- as.numeric(VARNAME_2_inf$X)
unique(VARNAME_2_inf$X)
result_df$ID <- as.numeric(result_df$ID)
unique(result_df$ID)

setdiff(unique(VARNAME_2_inf$X),unique(result_df$ID))

Weather_DATA<- left_join(result_df, VARNAME_2_inf,by=c('ID'='X'))

unique(paste(Weather_DATA$VARNAME_2, Weather_DATA$ID, sep = "_"))

unique(paste(Weather_DATA$VARNAME_2, Weather_DATA$NAME_1,Weather_DATA$NAME_2, sep = "_"))


MDR_2 <- sf::st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Raw_data/MDR_2/VMD_districts_2022.shp")

DAT_MDR_2<- left_join(Weather_DATA,MDR_2,by=c("VARNAME_2"="VARNAME_2","NAME_1" ="NAME_1" ,"NAME_2"="NAME_2" ))

library(sf)

Full_data<- st_drop_geometry(DAT_MDR_2)

filtered_data <- Full_data[Full_data$Year >= 2004, ]

dim(filtered_data)
names(filtered_data)

filtered_data<-filtered_data[,-c(13,17,18,19,20,21,22,23,24,25,26,27,16)]
names(filtered_data)

# Assuming your data frame is named filtered_data.
filtered_data_summarized <- filtered_data %>%  group_by(NAME_1 ,NAME_2, Year, Month)%>%
  mutate(
       NAME_2 = 
      ifelse(
        ( NAME_2 %in% c("Mỏ Cày Nam", "Mỏ Cày Bắc")) & NAME_1 %in% "Bến Tre",
       "Mỏ Cày",
        ifelse(
          ( NAME_2 %in%c( "Cờ Đỏ","Thới Lai")) & NAME_1 %in% "Cần Thơ",
          "Cờ Đỏ",
          ifelse(
            ( NAME_2 %in% c("Hồng Ngự (Thành phố)", "Hồng Ngự"))  & NAME_1 %in% "Đồng Tháp",
           "Hồng Ngự",
            ifelse(
              ( NAME_2 %in% c("Ngã Bảy", "Phụng Hiệp")) & NAME_1 %in% "Hậu Giang",
              "Ngã Bảy",
              ifelse(
                ( NAME_2 %in% c("Long Mỹ", "Long Mỹ (Thị xã)")) & NAME_1 %in% "Hậu Giang" ,
                "Long Mỹ",
                ifelse(
                  ( NAME_2 %in%  c("An Minh", "Vĩnh Thuận", "U Minh Thượng" ,"An Biên")) & NAME_1%in% "Kiên Giang",
                  "An Minh",
                  ifelse(
                    ( NAME_2 %in% c("Kiên Lương", "Giang Thành")) & NAME_1%in% "Kiên Giang",
                    "Kiên Lương",
                    ifelse(
                      ( NAME_2 %in% c("Mộc Hóa", "Kiến Tường")) & NAME_1 %in% "Long An",
                      "Mộc Hóa",
                      ifelse(
                        ( NAME_2 %in% c("Long Phú", "Mỹ Xuyên", "Trần Đề" )) & NAME_1 %in% "Sóc Trăng",
                        "Long Phú",
                        ifelse(
                        ( NAME_2 %in%c("Châu Thành", "Mỹ Tú" )) & NAME_1 %in% "Sóc Trăng",
                        "Mỹ Tú",
                        ifelse(
                          ( NAME_2 %in% c("Cai Lậy", "Cai Lậy (Thị xã)" ) )& NAME_1 %in% "Tiền Giang",
                          "Cai Lậy",
                          ifelse(
                            ( NAME_2 %in% c("Gò Công Đông", "Gò Công Tây", "Tân Phú Đông" )) & NAME_1 %in% "Tiền Giang" ,
                            "Gò Công Đông",
                            ifelse(
                              ( NAME_2 %in% c("Duyên Hải", "Duyên Hải (Thị xã)","Trà Cú")) & NAME_1 %in% "Trà Vinh" ,
                              "Trà Cú",
                              ifelse(
                                ( NAME_2 %in%  c("Bình Minh", "Bình Tân") ) & NAME_1 %in% "Vĩnh Long",
                               "Bình Minh",
                               ifelse(
      NAME_2 %in% c("Vĩnh Lợi", "Hoà Bình") & NAME_1 %in% "Bạc Liêu",
      "Vĩnh Lợi",  NAME_2
                              )
                            )
                          )
                        )
                      )
                    )
                  )
                )
              )
            )
          )
        )
      )
    )
  ) )%>%
 
  summarise(ws_avg = mean(ws_avg), ws_max = mean(ws_max), ws_min = mean(ws_min), t2m_avg = mean(t2m_avg), t2m_max = mean(t2m_max), t2m_min = mean(t2m_min), rh_avg = mean(rh_avg), rh_max = mean(rh_max),rh_min = mean(rh_min),tp_accum = mean(tp_accum)) %>%
  ungroup()


dim(filtered_data_summarized )


##Check the dimension to be similar to the dengue cases.
filtered_data_summarized_without_2023_year <- filtered_data_summarized %>%
  filter(!(Year == 2023))

MDR_NEW <- st_read(dsn = "C:/Users/uqwareed/OneDrive - The University of Queensland/Two months_predictions/Ensemble_model_monthly_data/Data/shapefiles/MDR_NEW_Boundaries_Final.shp") 

dim(filtered_data_summarized_without_2023_year )
MDR_Dengue_map_pre <- left_join(MDR_NEW, filtered_data_summarized, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))

dim(MDR_Dengue_map_pre)

selected_columns <- c(
   "NAME_1", "NAME_2", "TYPE_2","Year","Month","VARNAME",
  "ENGTYPE",  "ws_avg", "ws_max", "ws_min", "t2m_avg",
  "t2m_max", "t2m_min", "rh_avg", "rh_max", "rh_min", "tp_accum"
)

# Create the new data frame
D <- MDR_Dengue_map_pre %>%
  dplyr::select(all_of(selected_columns))

D1<- st_drop_geometry(D)
#write.csv(D1,'Weather_Data_with_New_Boundaries_2004_2023.csv')
#write.csv(filtered_data_summarized,'population_2004_and_more.csv')


 matching_district <- (MDR_NEW$NAME_2 %in% filtered_data_summarized$NAME_2 )
  matching_district 

# Display the matching values
matched_names <- unique(MDR_NEW$NAME_2[matching_district])

matched_names

Non_matching<- unique(MDR_NEW$NAME_2[!matching_district])


Non_matching
# Display the non-matching values
# Assuming matching_district is a logical vector
indices_not_matching <- which(!matching_district)

# Display the indices or positions where matching_district is FALSE
print(indices_not_matching)



MDR_weather_map <- left_join(MDR_NEW, filtered_data_summarized_without_2023_year, by = c("NAME_1" = "NAME_1" ,"NAME_2"="NAME_2"))


MDR_weather_map$NAME_En<- toupper(MDR_weather_map$NAME_En)
MDR_weather_map$VARNAME<- toupper(MDR_weather_map$VARNAME)

Final_set_dengue_weather<- inner_join(MDR_Dengue_map_Final, MDR_weather_map, by = c("province"  = "NAME_En" ,"district"="VARNAME","year"="Year","month"="Month","ID"="ID","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))


dim(Final_set_dengue_weather)

# Assuming your existing data frame is named existing_data_frame
# Assuming you want to select specific columns based on the provided names

# Columns to select
selected_columns <- c(
   "province", "district", "year", "month", "Dengue", "SevereDHF",
   "NAME_1", "NAME_2", "TYPE_2",
  "ENGTYPE",  "ws_avg", "ws_max", "ws_min", "t2m_avg",
  "t2m_max", "t2m_min", "rh_avg", "rh_max", "rh_min", "tp_accum"
)


# Create the new data frame
Final_subset_ERA5_weather <- Final_set_dengue_weather %>%
  dplyr::select(all_of(selected_columns))

##join it with pouopulstion data 

Final_ensemble_data<- inner_join(Final_set_subset_ERA5_weather, Final_subset_dengue_pop, by = c("province"  = "province","Dengue"="Dengue", "SevereDHF"= "SevereDHF" ,"district"="district","year"="year","month"="month","NAME_1"="NAME_1","NAME_2"="NAME_2","TYPE_2"="TYPE_2","ENGTYPE"="ENGTYPE"))

```

Formate the data for analysis and use lag



